\section{Introduction}

Machine-learning models are increasingly trained and deployed through
outsourced and cloud-based pipelines. Machine learning as a service
(MLaaS) allows clients with limited computational resources or machine-learning
expertise to outsource training and inference tasks to external providers
\cite{MLaaS15, machine_learning_survey20, VeriML21, privacy-mlaas23}. This paradigm is attractive in
settings such as risk scoring, healthcare analytics, fraud detection, and
financial decision making, where predictive models are often trained or served
over sensitive tabular data \cite{xgboost16,grinsztajn2022tree,shwartz2022tabular}.
However, outsourcing also creates an integrity problem: a client may receive a
trained model or a prediction from an external provider, but
cannot directly tell whether it was produced by the claimed training pipeline on
the claimed data source. Prior work on verifiable MLaaS and proof-of-learning
has therefore emphasized the need to certify that outsourced training or
inference was executed correctly, rather than replaced by a shortcut, a
partially trained model, or an improperly executed computation
\cite{VeriML21,proof_of_learning21,zkdnn24}.

At the same time, full transparency is often unacceptable. Training data may
contain sensitive user, medical, or financial records, while the learned model
itself may encode proprietary information or leak information about the
training data \cite{privacy-mlaas23,tramer2016stealing}. Model extraction
attacks further show that even black-box prediction APIs can compromise model
confidentiality, especially for models offered through cloud prediction
services \cite{tramer2016stealing}. This creates a tension between
verifiability and confidentiality: the verifier wants assurance that the model
was trained correctly, whereas the prover may need to keep the data, model
parameters, and intermediate training states private. Zero-knowledge proofs of
training address this tension by allowing a prover to convince a verifier that a
committed model is the output of a public training algorithm on a committed
dataset, while revealing nothing beyond the validity of this statement
\cite{zkdnn24,zkml_survey}.

In this work, we study zero-knowledge proofs for gradient boosted decision
trees (GBDTs). GBDTs, and especially implementations such as
XGBoost~\cite{xgboost16}, are among the strongest model families for tabular
data and are widely used in high-stakes decision pipelines built on structured
records, such as clinical risk prediction, financial risk assessment, and other
screening or allocation tasks. These applications are typically driven by
medium-sized tabular datasets with heterogeneous numeric and categorical
features, missing values, and irregular decision boundaries, a regime in which
tree-based models remain especially strong and often outperform or rival deep
tabular models in practice~\cite{shwartz2022tabular,grinsztajn2022tree,gorishniy2021revisiting}.
In such domains, model outputs can affect access to treatment, credit, and
other consequential opportunities, so it is important not only that the model
be accurate, but also that its training and predictions be auditable,
verifiable, and contestable~\cite{lima2022conflict,shahu2026faster}.

A GBDT model is an additive ensemble of regression
trees: each round trains new trees to fit the residual errors of the current
model, and the final prediction is obtained by aggregating the outputs of all
trees. This sequential structure makes training certification substantially
harder than inference certification. To prove inference, it is enough to show
that each input follows a valid path in each committed tree and that the
returned scores are aggregated correctly. To prove training, the prover must
also show that every split, every leaf value, and every boosting update is
consistent with the committed dataset and the public training rule.

Many prior zero-knowledge works on tree-based models focused on
inference, accuracy testing, or auditing properties of already trained trees.
zkDT~\cite{zkdt20} proves decision-tree predictions and accuracy claims without
revealing the tree, and later works improve committed-tree representations and
lookup-based tree inference~\cite{cq++24}. Confidential-PROFITT~\cite{confidential-profitt22}
certifies fairness properties of decision-tree training while keeping both the
model and training data confidential. These works provide important tools for
tree-model verification, but they do not give a specialized proof system for
the full training semantics of gradient boosting.

The closest line of work is zero-knowledge proof of training for tree
ensembles. Sparrow~\cite{sparrow24} gives a space-efficient zkSNARK for
data-parallel circuits and applies it to random-forest training and prediction.
Its key idea is to avoid re-running training inside the proof: instead, the
prover proves that a certification algorithm accepts the committed forest and
dataset. ZKBoost~\cite{zkboost26} extends this certification approach to XGBoost.
It defines a deterministic fixed-point version of XGBoost, builds a
certification algorithm for the resulting trainer, and then proves this
certifier using a general-purpose ZKP backend. These works demonstrate that
proofs of training for tree ensembles are possible, but their proof systems
still inherit the cost of proving the computation of large certification circuits.

A natural alternative is to avoid proving the certification computation as a
generic circuit and instead prove the algebraic relations that characterize
GBDT training directly. However, this approach creates a different bottleneck.
GBDT training contains many repeated constraints of two different kinds. The
first kind consists of \emph{low-degree algebraic constraints}: relations that
can be written as low-degree polynomial identities over the field, such as
score updates, prefix sums, booleanity checks, and
zero checks. The second kind consists of
\emph{non-native predicates}: checks whose intended semantics depend on
structure outside the field algebra, such as table membership, order, range,
integer divisibility, rounding, and maximum selection.
These constraints also do not naturally live on one uniform domain,
since their inputs have different sizes.
Different tensors have different shapes, and their polynomial encodings are
committed over different subgroups. Proving each relation separately would
introduce a large number of auxiliary commitments and openings, while ordinary
same-domain batching does not apply directly.

We present \textsc{Terrae}, a zero-knowledge proof system for GBDT
training and inference based on KZG polynomial commitments. \textsc{Terrae}
proves that a committed forest is obtained by executing a public GBDT
trainer on a committed dataset.
The proof of \textsc{Terrae}
follows the structure of histogram-based tree training verification~\cite{sparrow24,zkboost26}:
it proves that samples in the dataset
are routed to the correct leaves, that sample-wise residuals are correctly
converted into per-node and per-bin histograms, that the selected split at each
internal node is optimal according to the public splitting rule, that leaf
values are computed from the committed residuals, and that boosting states are
updated consistently across rounds.
However, unlike previous approaches
that compile these certification procedures into
large circuits,
\textsc{Terrae} represents the training trace as
committed tensors and proves the relation among them directly.

Besides handling both low-degree algebraic constraints and non-native
predicates,
the main technical challenge is batching. Once the training semantics are
decomposed into algebraic constraints, the prover faces many similar constraints
over different domains. \textsc{Terrae} introduces two batching tools to handle
this heterogeneity. First, \emph{domain-lifting batching} combines low-degree
algebraic constraints, such as zero-check and sum-check claims, across
different domains. It
lifts all residual polynomials to a common domain and then applies a
single random linear combination. Second, \emph{interleaving batching} combines
same-form non-native predicates, including lookup-style checks, by virtually
interleaving committed vectors of different lengths. 
Any evaluation of the resulting polynomial can be derived efficiently from existing polynomial commitments and public sparse polynomials, 
and the batching does not introduce any new commitments to auxiliary witnesses.

We also design a dedicated histogram proof for GBDT training. Prior
certification-based approaches treat histogram construction as a RAM-like
update procedure~\cite{sparrow24,zkboost26},
where each sample updates the memory slot indexed by its leaf, feature,
and bin, by reading the value from and writing the updated value back to the same location.
In contrast, \textsc{Terrae} exploits the fact that histogram updates
are fixed additive updates with a much simpler access pattern than a generic RAM.
It proves the final histogram by binding each bucket
to a power of random challenge and checking a weighted-sum identity between the committed
histogram and the sample-wise update stream. This turns histogram verification
from a sequence of authenticated memory transitions into a single randomized
aggregate identity. It avoids proving a general RAM
computation and gives a proof primitive tailored to the histogram structure
that appears throughout tree training.

Our main contributions are as follows:

\begin{itemize}
    \item We present \textsc{Terrae}, a proof system that
    proves the training and inference of GBDTs in zero knowledge.
    \textsc{Terrae} avoids compiling each training or inference
    certification procedure into
    separate circuits, and instead proves the algebraic relations directly.

    \item We introduce domain-lifting batching for low-degree algebraic
    constraints and interleaving batching for non-native predicates. These techniques
    batch constraints that arise over differently sized domains and reduce many
    commitment-opening tasks to a small number of batched claims, without
    committing to extra witnesses.

    \item We propose a histogram proof that proves the correctness of converting
    sample-wise training data and residuals into per-node, per-feature, and
    per-bin histograms. The proof exploits the additive structure of histogram
    construction and avoids general RAM consistency checking.

    \item We implement \textsc{Terrae} and evaluate it against a
    baseline following the design of prior proof-of-training
    systems for tree ensembles. The results show that directly proving the
    algebraic structure of GBDT training \textcolor{red}{significantly} reduces proof-generation
    time, while adding only \textcolor{red}{modest} proof-size overhead.
\end{itemize}

The rest of the paper is organized as follows. \Cref{sec:background}
introduces the notation, background, and plaintext training setting.
\Cref{sec:threat-model} states the threat model and security goal.
\Cref{sec:technical-overview} summarizes the main technical ideas, and
\Cref{sec:methodology} describes the proof methodology and algebraic checks.
\Cref{sec:evaluation} reports the experimental results.
\Cref{sec:conclusion} concludes the technical presentation.
We defer related work, the formal batching tools, and the full ZK-GBDT protocol
to \Cref{sec:related,sec:tools,sec:zkgbdt-protocol}, and defer extensions to
random forests and XGBoost-style training, as well as constructions of selected
subprotocols, to the remaining appendices.
