\section{Technical Overview}\label{sec:technical-overview}

In this section, we introduce four ingredients: 
how to commit the forest and prove inferences,
how to prove the correctness of training,
how to prove the correctness of histogram,
and how to batch claims to reduce the overall proof size.

\subsection{Forest Representation}

We adopt the hyperbox forest encoding of \cite{cq++24}. Each node carries a
feature-wise half-open interval box, and a sample reaches a leaf exactly when
its quantized feature vector lies inside that box. 
We assume there are $N_{\mathrm{tot}}$ nodes, and
all committed objects share
one global node namespace, so node identifiers are
$u \in \{0,\ldots,N_{\mathrm{tot}}-1\}$ everywhere. Node $0$ is a dummy node, and
all of its committed attributes are fixed to the dummy value $0$. When a node identifier labels
a committed row, the corresponding physical row in the row-major encoding is
$u$. The committed
forest consists of the commitments of:
\begin{itemize}
  \item left and right child index vectors
  $\mathbf{lc},\mathbf{rc}\in\{0,1,\ldots,N_{\mathrm{tot}}-1\}^{N_{\mathrm{tot}}}$,
  whose internal-node entries record the node identifiers of the left and
  right child of each parent node, while leaf and padding entries are zero;
  \item lower and upper node bounds
  $\mathbf{N}^{\leftarrow},\mathbf{N}^{\rightarrow}\in[B+1]^{N_{\mathrm{tot}}\times d}$,
  where row $u$ stores the $d$-dimensional half-open box of node $u$ and every
  leaf row describes one accepting region of the tree;
  \item a feature-selection matrix
  $\mathbf{E}\in\{0,1\}^{N_{\mathrm{tot}}\times d}$,
  whose internal-node rows are one-hot and indicate which feature is split at
  each node, while leaf rows are zero;
  \item a threshold-index vector
  $\boldsymbol{\Theta}\in\{0,1,\ldots,B\}^{N_{\mathrm{tot}}}$,
  whose internal-node entries record the selected threshold bin and whose leaf rows are zero;
  \item a tree-index vector $\mathbf{tree}\in[TK+1]^{N_{\mathrm{tot}}}$, where trees are ranked from $1$ to $TK$
  except dummy node $0$, and a
  root indicator $\mathbf{root}\in\{0,1\}^{N_{\mathrm{tot}}}$;
  \item a leaf-score vector $\mathbf{score}\in\mathbb{F}^{N_{\mathrm{tot}}}$,
  whose leaf entries store the scalar contribution returned by that tree and
  whose internal-node entries are zero;
  \item a boolean vector $\mathbf{is\_leaf}\in\{0,1\}^{N_{\mathrm{tot}}}$, 
  where $\mathbf{is\_leaf}[u]=1$ if and only if $u$ is a leaf node.
\end{itemize}

We assume there are $N_{\mathrm{int}}$ internal nodes and $N_{\mathrm{leaf}}$ leaf nodes
in the forest, where internal nodes are indexed by $\{1,\ldots,N_{\mathrm{int}}\}$ and leaf nodes are indexed by
$\{N_{\mathrm{int}}+1, \dots, N_{\mathrm{tot}}-1\}$.

Note that $\mathbf{N}^{\leftarrow}$, $\mathbf{N}^{\rightarrow}$, $\mathbf{lc}$
and $\mathbf{rc}$ together fix the entire forest.
Other vectors/matrices are auxiliary to the proof
of training and inference.

\Cref{fig:forest-encoding-inference} shows the intuition behind
the forest encoding: the left panel illustrates the structure of the forest,
while the right panel presents the committed objects.

\input{figures/forest_encoding_inference}

\subsection{Proof of Inference}

To prove the correctness of inference over batched input $\mathbf{I}\in\mathbb{F}^{b\times d}$,
the prover commits $\mathbf{U}^{\leftarrow}, \mathbf{U}^{\rightarrow}\in \mathbb{F}^{TKb\times d}$,
$\mathbf{sc}\in\mathbb{F}^{TKb}$,
and utilize table lookup to prove
\begin{align}
  \mathbf{1} \| \mathbf{ind\_tree}  \| \mathbf{U} ^{\leftarrow } \| \mathbf{U} ^{\rightarrow } \| \mathbf{sc} \subseteq
  \mathbf{is\_leaf} \| \mathbf{tree} \| \mathbf{N} ^{\leftarrow }\| \mathbf{N} ^{\rightarrow }\| \mathbf{score},\label{eq:inference-lookup-1}
\end{align}
where notation ``$\subseteq$'' means that each row of the left matrix equals some row of the right matrix, and
\[
\mathbf{ind\_tree} \coloneqq (\underbrace{0, \dots, 0}_b, \underbrace{1, \dots, 1}_{b}, \dots, \underbrace{TK-1, \dots, TK-1}_b)^\top.
\]
As a sanity check,
the LHS of \Cref{eq:inference-lookup-1} has shape $TKb\times (2d+3)$,
and the RHS has shape $N_{\mathrm{tot}}\times (2d+3)$, so the table lookup is well-defined.
To prove this lookup, we pad each column vector to a $d$ column matrix,
randomly combine the matrices with a verifier challenge,
and utilize the table lookup of~\cite{cq++24} to prove the lookup
with (online) prover time $O(TKb)$.

This lookup argument guaranttes that each row of 
$\mathbf{U}^{\leftarrow}, \mathbf{U}^{\rightarrow}$
and $\mathbf{sc}$ are aligned for the same leaf.
By further proving
\begin{align}
  \mathbf{U} ^{\leftarrow }_{\{T\times K\times b\times d\}}[t, k, i, j] \leq \mathbf{I}[i,j] \leq \mathbf{U}_{\{T\times K\times b\times d\}} ^{\rightarrow }[t, k, i, j], \label{eq:inference-interval}
\end{align}
the prover proves that the leaf reached by each sample is correct.
What remains is to prove the result of inference from $\mathbf{sc}$,
where $\mathbf{sc}_{\{T\times K\times b\}}[t,k,i]$ represents the score sample $i$
received from the tree trained at round $t$ for class $k$.
This includes proving the sum of $\mathbf{sc}_{\{T\times K\times b\}}$
along the second dimension and proving the location of the maximum.

\subsection{Proof of Forest Well-formedness}\label{subsec:proof of well-formedness}

First, the prover proves that the committed forest is well-formed, by showing
\begin{enumerate}
  \item The left and right children share the same tree index with the parent,
    i.e. $\mathbf{tree}[\mathbf{lc}[u]]=\mathbf{tree}[\mathbf{rc}[u]]=\mathbf{tree}[u]$ for all internal nodes $u$.
  \item Leaves have no children, i.e. \begin{align}
    \mathbf{lc}\odot \mathbf{is\_leaf}=\mathbf{0}\land \mathbf{rc}\odot \mathbf{is\_leaf}=\mathbf{0},\label{eq:leaf-no-child}
  \end{align}
  where ``$\odot$'' is the Hadamard product.
  \item The hyper-box of any root is full, i.e.
  \begin{align}
    \mathbf{Root}\odot \mathbf{N}^{\leftarrow}=\mathbf{0}\land \mathbf{Root}\odot \mathbf{N}^{\rightarrow}=(B+1)\mathbf{Root},\label{eq:root-full-box}
  \end{align}
  where $\mathbf{Root} = (\mathbf{root}\|\cdots\|\mathbf{root})$.
  \item The hyper-box of left and right children are consistent with the parent and the split, i.e. for each internal node $u$ and feature $j$,
    \begin{align}
    \mathbf{N}^{\leftarrow}[u, j] - \mathbf{N}^{\leftarrow}[\mathbf{lc}[u], j] &= 0\land\label{eq:left-child-inherits-lowerbound}\\
    \mathbf{N}^{\rightarrow}[u, j] - \mathbf{N}^{\rightarrow}[\mathbf{rc}[u], j] &= 0\land\label{eq:right-child-inherits-upperbound}\\
    \mathbf{E}[u, j](\mathbf{N}^{\rightarrow}[\mathbf{lc}[u], j] - \mathbf{N}^{\leftarrow}[\mathbf{rc}[u], j]) &= 0\land\label{eq:split-child-consistency}\\
    (1-\mathbf{E}[u, j])(\mathbf{N}^{\leftarrow}[u, j] - \mathbf{N}^{\leftarrow}[\mathbf{rc}[u], j]) &= 0\land\label{eq:non-split-child-inherits-1}\\
    (1-\mathbf{E}[u, j])(\mathbf{N}^{\rightarrow}[u, j] - \mathbf{N}^{\rightarrow}[\mathbf{lc}[u], j]) &= 0.\label{eq:non-split-child-inherits-2}
    \end{align}
    \Cref{eq:left-child-inherits-lowerbound} and \ref{eq:right-child-inherits-upperbound} ensure that 
    left child inherits the lower bound of the parent and right child inherits the upper bound.
    \Cref{eq:split-child-consistency} ensures that for the split feature, 
    the left child's upper bound meets the right child's lower bound.
    \Cref{eq:non-split-child-inherits-1} and \ref{eq:non-split-child-inherits-2} together ensure that 
    for non-split features, the child hyper-box inherits the parent hyper-box.
  \item $\mathbf{E}$ is boolean and one-hot on internal nodes and zero on leaves, i.e. for each node $u$,
    \begin{align}
    \sum_{j=1}^d \mathbf{E}[u, j] - \mathbf{is\_leaf}[u] + 1 &= 0,\\
    \mathbf{E}\odot(\mathbf{E}-\mathbf{1}) &= \mathbf{0}.
    \end{align}
  \item $\mathbf{score}$ is zero on internal nodes, i.e.
    \begin{align}
    \mathbf{score}\odot (\mathbf{1}-\mathbf{is\_leaf}) = \mathbf{0}.
    \end{align}
  \item There are in total $TK$ trees, i.e.
  \begin{align}
    \sum_{u=0}^{N_{\mathrm{tot}}-1} \mathbf{root}[u] = TK.
  \end{align}
  \item Node $0$ is dummy, i.e.
    \begin{align}
    \mathbf{lc}[0] = \mathbf{rc}[0] &= 0,\\
    \forall j\in [d].\ \mathbf{N}^{\leftarrow}[0, j] = \mathbf{N}^{\rightarrow}[0, j] &= 0,\\
    \mathbf{tree}[0] = \mathbf{root}[0] = \mathbf{is\_leaf}[0] &= 0.
    \end{align}
  \end{enumerate}

Note that since for any node $u$, its hyper-box is full-range if and only if
  $\mathbf{root}[u]=1$,
  any node $v$ marked by $\mathbf{root}[v]=1$ must be a true root.

The above involves essentially two kinds of techniques:
  proving routing, and proving sum.
  To prove $\mathbf{tree}[\mathbf{lc}[u]]=\mathbf{tree}[u]$ for all internal nodes $u$, the prover first
  commits $\mathbf{tree\_lc}$ where
  \[\mathbf{tree\_lc}[u] = \mathbf{tree}[\mathbf{lc}[u]]\]
  for all internal nodes $u$.
  It then proves
  \[
    \mathbf{lc}\|\mathbf{tree\_lc} \subseteq \left(\begin{matrix}
    0\\ 1\\ \vdots\\ N_{\mathrm{tot}}-1
    \end{matrix}\right)\|\mathbf{tree},
  \]
  alongside $\mathbf{is\_leaf}\odot(\mathbf{tree\_lc} - \mathbf{tree}) = \mathbf{0}$.

For sum over vector, we utilize the Aurora's sum-check protocol~\cite{aurora19}.
  To prove sum along one dimension of the matrix,
  we adopt a Quarks-style proof~\cite{quarks20},
  where the prover commits the prefix sum and total sum, 
  and proves the consistency between them and the original matrix.

\subsection{Proof of Training}

The training proof reuses the same committed forest. 
The prover first commits the dataset $\mathbf{D}\in [B]^{M\times d}$ and 
its one-hot label $\mathbf{Y}\in \{0, 1\}^{K\times M}$.
The prover then commits leaves reached by each sample in the dataset,
by committing
\[
\mathbf{Ind} \in
\{N_{\mathrm{int}}+1,\ldots,N_{\mathrm{tot}}-1\}^{T\times K\times M},
\]
where $\mathbf{Ind}[t, k, r]$ 
is the index of the leaf reached by sample
$r$ in the tree trained for class $k$ at round $t$.

Likewise, the prover also commits the pseudo-residuals
$\mathbf{G}\in \mathbb{F}^{T\times K\times M}$ and the current scores
$\mathbf{S}\in \mathbb{F}^{T\times K\times M}$.
For round $t$, class $k$, and sample $r$,
$\mathbf{G}[t,k,r]$ represents the pseudo-residual target of sample $r$ that
should be approximated by the regression tree for class $k$ at round $t$, and
$\mathbf{S}[t,k,r]$ represents the cumulative score of sample $r$ for class
$k$ after round $t$ has been added.
Initially, $\mathbf{G}[0,k,r]=\mathbf{Y}[k,r]$ for all $k$ and $r$.

Logically, the proof of training sequentially proves that:
\begin{enumerate}
  \item $\mathbf{Ind}$ correctly commits the leaf reached by each sample in the dataset.
    This is proved in a manner similar to proving batched inference.
  \item $\mathbf{G}[0, \cdot, \cdot]$ is set correctly from $\mathbf{Y}$,
    namely $\mathbf{G}[0,k,r]=\mathbf{Y}[k,r]$.
  \item The tree trained at each round $t$ is the optimal regression tree that fits $\mathbf{G}[t, \cdot, \cdot]$ on the dataset,
    i.e. each node split is optimal and the score is assigned to the leaves correctly. This is discussed in more detail in the next subsections.
  \item For each round $t$, the score $\mathbf{S}[t, \cdot, \cdot]$ is updated correctly
    from the previous score state and the scores each sample obtained from the trees trained at round $t$.
  \item For each round $t<T-1$, $\mathbf{G}[t+1, \cdot, \cdot]$ is updated correctly from $\mathbf{Y}$ and $\mathbf{S}[t, \cdot, \cdot]$.
    This involves proving softmax over the second dimension of $\mathbf{S}$.
\end{enumerate}

Together, these conditions imply that the committed objects encode a valid
plaintext GBDT training trace across all rounds.

\subsection{Proving the Correctness of Node Split}

Consider one node $u$, one feature $j$, and one threshold bin $b$. Let
$I_u \subseteq [M]$ be the samples reaching node $u$, and let
$L_{u,j,b},R_{u,j,b}$ be the induced left and right partitions. If each sample
$r\in I_u$ carries target value $g_r$, then the residual-sum-of-squares
objective is
\[
\operatorname{RSS}(u,j,b)
=
\sum_{r\in L_{u,j,b}}(g_r-\bar{g}_L)^2
+
\sum_{r\in R_{u,j,b}}(g_r-\bar{g}_R)^2.
\]
Expanding the squares gives
\begin{equation}\label{eq:rss-rewrite}
\operatorname{RSS}(u,j,b)
=
\sum_{r\in I_u} g_r^2
-
\left(
\frac{\left(\sum_{r\in L_{u,j,b}} g_r\right)^2}{|L_{u,j,b}|}
+
\frac{\left(\sum_{r\in R_{u,j,b}} g_r\right)^2}{|R_{u,j,b}|}
\right).
\end{equation}
For fixed node $u$, the first term in \eqref{eq:rss-rewrite} is independent of
the candidate split. Therefore minimizing $\operatorname{RSS}(u,j,b)$ is
equivalent to maximizing
\begin{equation}\label{eq:split-objective}
\Phi(u,j,b)
\coloneqq
\frac{\left(\sum_{r\in L_{u,j,b}} g_r\right)^2}{|L_{u,j,b}|}
+
\frac{\left(\sum_{r\in R_{u,j,b}} g_r\right)^2}{|R_{u,j,b}|},
\end{equation}
subject to the validity condition that both children are nonempty.

If we denote $\mathbf{C}[i, j, k]$ the number of samples falling to node $i$ with $d$-th feature being $k$,
  and $\mathbf{V}[i, j, k]$ the sum of residuals for samples falling to node $i$ with $d$-th feature being $k$,
  then we can rewrite \eqref{eq:split-objective} as
\begin{equation}
\Phi(u,j,b)
\coloneqq
\frac{\left(\sum_{k=0}^{b-1} \mathbf{V}[u, j, k]\right)^2}{\sum_{k=0}^{b-1} \mathbf{C}[u, j, k]}
+
\frac{\left(\sum_{k=b}^{B-1} \mathbf{V}[u, j, k]\right)^2}{\sum_{k=b}^{B-1} \mathbf{C}[u, j, k]}.
\end{equation}

This rewriting indicates that \emph{histograms} are sufficient proof objects. 
Once $\mathbf{C}$ and $\mathbf{V}$ are available, by proving the prefix and suffix sums of them,
the prover can prove the split gain $\Phi(u,j,b)$ for all candidate splits in one shot.

\subsection{Proving Histogram}

Given $\mathbf{Ind}$ and $\mathbf{G}$,
  the prover needs to prove the correctness of $\mathbf{C}$ and $\mathbf{V}$.
  Following the approach of \cite{sparrow24,zkboost26},
  we first construct the leaf-wise histogram $\mathbf{C}_{\mathrm{leaf}}$ and $\mathbf{V}_{\mathrm{leaf}}$,
  where $\mathbf{C}_{\mathrm{leaf}}[i, j, k]$ and $\mathbf{V}_{\mathrm{leaf}}[i, j, k]$ are 
  the number of samples and sum of residuals for node $i$ (which is a leaf), feature $j$, and feature value $k$.
  The prover then proves $\mathbf{C}$ and $\mathbf{V}$ by showing that for all nodes,
  \begin{align}
  \mathbf{C}[i, j, k] =&\ \mathbf{C}_{\mathrm{leaf}}[l, j, k]\cdot \mathbf{is\_leaf}[l] +\nonumber\\
    &(\mathbf{C}[\mathbf{lc}[i], j, k] + \mathbf{C}[\mathbf{rc}[i], j, k])\cdot (1- \mathbf{is\_leaf}[l]),\\
  \mathbf{V}[i, j, k] =&\ \mathbf{V}_{\mathrm{leaf}}[l, j, k]\cdot \mathbf{is\_leaf}[l] +\nonumber\\
    &(\mathbf{V}[\mathbf{lc}[i], j, k] + \mathbf{V}[\mathbf{rc}[i], j, k])\cdot (1- \mathbf{is\_leaf}[l]),
  \end{align}
  where again, the prover proves the routing with lookup.

Hence, the prover first needs to prove that
  \begin{align}
    \mathbf{C}_{\mathrm{leaf}}[i, j, k] =&\ \sum_{r: \mathbf{Ind}[t, k, r]=i \land \mathbf{D}[r, j]=k} 1,\\
    \mathbf{V}_{\mathrm{leaf}}[i, j, k] =&\ \sum_{r: \mathbf{Ind}[t, k, r]=i \land \mathbf{D}[r, j]=k} \mathbf{G}[t, k, r].
  \end{align}
  Sparrow~\cite{sparrow24} and ZKBoost~\cite{zkboost26} prove this by modeling it as an RAM argument.
  That is, consider an RAM array $\mathbf{C}'$ and $\mathbf{V}'$, for each sample $r$, we update
  \[
  \begin{aligned}
    \mathbf{C}'[\mathbf{Ind}[t, k, r], j, \mathbf{D}[r, j]] &\gets \mathbf{C}'[\mathbf{Ind}[t, k, r], j, \mathbf{D}[r, j]] + 1,\\
    \mathbf{V}'[\mathbf{Ind}[t, k, r], j, \mathbf{D}[r, j]] &\gets \mathbf{V}'[\mathbf{Ind}[t, k, r], j, \mathbf{D}[r, j]] + \mathbf{G}[t, k, r].
  \end{aligned}
  \]
  They then prove this by utilizing the RAM argument of Spartan~\cite{spartan20} and of~\cite{twoshuffleRAM24}, respectively.

However, we note that the problem of proving histogram is one
  that is much simpler than general RAM argument, since the update pattern is fixed and known in advance, and the update operation is just addition.
  Over the flattened tensors, we instead prove
  \begin{align}
    \sum_{i\in [TKM]} \mathbf{C}_{\{TKM\}}[i]\cdot \alpha^i &= \sum_{t, k, r, j} \alpha^{\mathsf{ind}(t, k, r, j)}\cdot 1,\\
    \sum_{i\in [TKM]} \mathbf{V}_{\{TKM\}}[i]\cdot \alpha^i &= \sum_{t, k, r, j} \alpha^{\mathsf{ind}(t, k, r, j)}\cdot \mathbf{G}[t, k, r],
  \end{align}
  where $\alpha$ is a random challenge and $\mathsf{ind}(t, k, r, j)$ is the flattened index of the tuple $(t, k, r, j)$,
  namely
  \[\mathsf{ind}(t, k, r, j) = \mathbf{Ind}[t, k, r]\cdot KM + j\cdot M + \mathbf{D}[r, j].\]
  Intuitively, we bind each $i$-th memory slot to an $\alpha^i$,
  and prove the correctness of the final state by proving the sum of all memory slots weighted by $\alpha^i$.
  
Hence in our construction, the prover
  first commits and proves a tensor encoding $\mathsf{ind}(t, k, r, j)$
  for all $(t, k, r, j)$,
  then utilizes lookup argument to prove the correctness of power-of-$\alpha$,
  and finally proves the statement with a Aurora-style sum-check protocol~\cite{aurora19}.

\subsection{Domain-Lifting and Interleaving Batching}

As can be already noticed, our construction repeatedly 
encounter checks of similar forms, e.g. in \Cref{subsec:proof of well-formedness},
we repeatedly require proof that a polynomial evaluates to zero
everywhere on a domain (zero-check), or repeatedly require lookup argument.
What makes the problem tricky is that these checks
happens over different domains, and a usual random linear
combination will not help.
In this subsection, we discuss the two approaches
we propose to solve this problem.

\paragraph{Domain-lifted batch.}
As stated, most linear constraints in the paper are first expressed over some evaluation
domain $\mathbb{H}_n$. After moving all terms to one side, we obtain a virtual
polynomial $F(X)$, and the constraint becomes a zero-check over that domain,
hence a divisibility claim of the form
\[
F(X)=Q(X)\nu_n(X),
\]
where $\nu_n(X)=X^n-1$ is the vanishing polynomial of $\mathbb{H}_n$.

The basic batching problem is therefore the following: suppose we have two such
claims, one over $\mathbb{H}_{n_1}$ and one over $\mathbb{H}_{n_2}$, with
$n_1<n_2$. Since our domain sizes are powers of two, the smaller domain is
nested inside the larger one. We can therefore lift the smaller claim to
$\mathbb{H}_{n_2}$ by multiplying its residual by
$\nu_{n_2}(X)/\nu_{n_1}(X)$:
\[
\widehat{F}_1(X):=F_1(X)\frac{\nu_{n_2}(X)}{\nu_{n_1}(X)}.
\]
Now both claims live over the same ambient domain $\mathbb{H}_{n_2}$, and the
verifier can batch them as
\[
\widehat{F}_1(X)+\eta F_2(X)=Q(X)\nu_{n_2}(X)
\]
for a random challenge $\eta$.

This is the basic intuition behind \emph{domain-lifting batching}: lift the
smaller residual to the larger domain, then batch there.

\paragraph{Interleaving batching.}
The second batching problem concerns same-form nonlinear checks that are
\emph{local}: validity at one position depends only on the witness values at
that position, not on any other entry of the vector. Typical examples in our
setting include lookup-style checks and fixed-point quantization checks with one
shared divisor and rounding rule.

The basic case is two vectors of the same length. Suppose
$\mathbf{f},\mathbf{g}\in\mathbb{F}^n$ must both satisfy the same local
constraint, say $\varphi(x)=0$ entrywise. Then we can interleave them into one
larger virtual vector
\[
(\mathbf{f}[0],\mathbf{g}[0],\mathbf{f}[1],\mathbf{g}[1],\ldots),
\]
and check the same local constraint once on that merged vector. In polynomial
form, if $f(X)$ and $g(X)$ are the encodings of $\mathbf{f}$ and $\mathbf{g}$
over $\mathbb{H}_n$, the merged object is
\[
\operatorname{Int}_n(f,g)(X)
:=
\operatorname{even}_n(X)\,f(X)
+ \operatorname{odd}_n(X)\,g(\omega_{2n}^{-1}X),
\]
where
\[
\operatorname{even}_n(X):=\frac{X^n+1}{2},
\qquad
\operatorname{odd}_n(X):=\frac{1-X^n}{2}.
\]
These public sparse selectors place $\mathbf{f}$ on the even positions and
$\mathbf{g}$ on the odd positions, so every position of the larger vector still
corresponds to one concrete original entry.

If the two vectors have different lengths, we first lift the smaller one to the
larger domain by the substitution $X\mapsto X^r$, exactly as in
\Cref{lem:row-replication}. After this alignment step, the same interleaving
idea applies.
The important point here is that the merged object is still virtual: it is assembled
from existing commitments, public sparse masks, and domain-lift substitutions,
so no new commitment is required.
