\section{Construction}\label{sec:construction}

\subsection{Compatibility Layer}

The protocol repeatedly needs to combine checks that live on different subgroup
sizes or have incompatible tensor layouts. We therefore start with a small
compatibility layer that explains how these objects are aligned before they are
used in lookup, histogram, or split-selection checks.

\begin{lemma}[Domain lifting]
\label{lem:domain-lifting}
Let
\[
1 \leq n_1 < n_2 < \cdots < n_m = N
\]
be powers of two. For each $i \in [m]$, let $F_i(X)$ be a virtual polynomial
and suppose the target claim is
\[
F_i(X)=Q_i(X)\nu_{n_i}(X)
\]
for some polynomial $Q_i(X)$. Define the lifted residual
\[
\widehat{F}_i(X):=
F_i(X)\frac{\nu_N(X)}{\nu_{n_i}(X)}.
\]
Then:
\begin{enumerate}
  \item $F_i(X)=Q_i(X)\nu_{n_i}(X)$ holds if and only if
  $\widehat{F}_i(X)$ is divisible by $\nu_N(X)$.
  \item For a verifier challenge $\eta$, all $m$ claims can be combined into
  the single identity
  \begin{equation}\label{eq:domain-lifted-batch}
  \sum_{i=1}^{m}\eta^{i-1}\widehat{F}_i(X)=Q(X)\nu_N(X).
  \end{equation}
  \item If at least one original claim is false, then
  \eqref{eq:domain-lifted-batch} holds for at most $m-1$ values of $\eta$.
\end{enumerate}
\end{lemma}

\begin{proof}
The first item is immediate from
\[
F_i(X)=Q_i(X)\nu_{n_i}(X)
\iff
F_i(X)\frac{\nu_N(X)}{\nu_{n_i}(X)} = Q_i(X)\nu_N(X).
\]
For soundness, once the lifted residuals are reduced modulo $\nu_N(X)$, the
left-hand side of \eqref{eq:domain-lifted-batch} becomes a polynomial in $\eta$
of degree at most $m-1$. If at least one original claim is false, this
polynomial is nonzero, so a uniformly random $\eta$ causes accidental
acceptance with probability at most $(m-1)/|\mathbb{F}|$.
\end{proof}

\begin{lemma}[Row replication]
\label{lem:row-replication}
Let $\mathbf{v}\in\mathbb{F}^m$ and let $v(X)$ be its encoding over
$\mathbb{H}_m$. Let $\mathbf{V}\in\mathbb{F}^{r\times m}$ be the matrix whose
every row equals $\mathbf{v}$. Then the encoding polynomial of $\mathbf{V}$ over
$\mathbb{H}_{rm}$ is
\[
V(X)=v(X^r).
\]
\end{lemma}

\begin{proof}
For every $i\in[r]$ and $j\in[m]$,
\[
V(\omega_{rm}^{im+j})
= \mathbf{V}[i][j]
= \mathbf{v}[j]
= v(\omega_m^j)
= v((\omega_{rm}^{im+j})^r),
\]
where the last equality uses $\omega_{rm}^r=\omega_m$. Since both sides have
degree less than $rm$ and agree on $\mathbb{H}_{rm}$, they are identical.
\end{proof}

\begin{definition}[Interleaving]
Let $\mathbf{a}^{(0)},\ldots,\mathbf{a}^{(k-1)}\in\mathbb{F}^n$. Their
interleaving is the vector
\[
\operatorname{Int}\!\left(
\mathbf{a}^{(0)},\ldots,\mathbf{a}^{(k-1)}
\right)\in\mathbb{F}^{kn}
\]
defined by
\[
\operatorname{Int}\!\left(
\mathbf{a}^{(0)},\ldots,\mathbf{a}^{(k-1)}
\right)[ik+t]
:=
\mathbf{a}^{(t)}[i]
\qquad
\forall i\in[n],\ t\in[k].
\]
\end{definition}

Interleaving is used when several pointwise nonlinear checks should be proved as
one relation on a shared domain. The most common example in our setting comes
from lookup-based subprotocols. A lookup proof often produces several
pointwise multiplicative constraints, such as reversal checks of the form
\[
h[i]\cdot (\gamma-f[i]) = c[i]
\qquad \forall i \in [n],
\]
where $f$, $h$, and $c$ are committed vectors and $\gamma$ is a verifier
challenge. If two such checks arise on different domain sizes, say
$n_1<n_2$, proving them separately would require separate openings and separate
quotient witnesses even though the pointwise pattern is identical.

Interleaving removes this duplication. We first align the shorter bundle to the
longer domain using row replication or domain lifting. We then form one larger
bundle by alternating entries from the two aligned bundles. If the two bundles
obey the same pointwise relation, the verifier now needs to check only one
relation on the interleaved object. If several relation types are present, we
interleave them together with public selector masks indicating which formula is
active at each position.

For example, suppose we must prove
\[
h_1[i](\gamma-f_1[i]) = c_1[i]
\quad \forall i\in[n_1]
\qquad\text{and}\qquad
h_2[j](\gamma-f_2[j]) = c_2[j]
\quad \forall j\in[n_2].
\]
After aligning the first triple $(h_1,f_1,c_1)$ to $\mathbb{H}_{n_2}$, we
interleave the aligned triples and obtain vectors
$\widehat{h},\widehat{f},\widehat{c}$ of length $2n_2$ satisfying
\[
\widehat{h}[z]\cdot(\gamma-\widehat{f}[z])=\widehat{c}[z]
\qquad \forall z \in [2n_2].
\]
This is why interleaving is particularly useful for lookup-style checks: it
turns many small nonlinear relations into one larger nonlinear relation whose
shape is easier to batch with the rest of the protocol.

\paragraph{Greedy interleaving scheduler.}
In practice, we apply interleaving to bundles of polynomials rather than to
single vectors. Each bundle contains all tensors that participate in one
pointwise nonlinear relation. The scheduler below repeatedly merges the two
smallest active bundles until only one remains.

\begin{quote}
\small
\noindent\textbf{GreedyInterleave}$(\mathcal{B})$\\
\textbf{Input:} a multiset of bundles $\mathcal{B}$, where each bundle stores
its committed tensors and its domain size.\\
\textbf{1.} Bucket all bundles by domain size.\\
\textbf{2.} While more than one bundle remains active: \\
\hspace*{1.5em}\textbf{a.} Remove two bundles with smallest domain sizes,
say $(\mathcal{U}_1,n_1)$ and $(\mathcal{U}_2,n_2)$ with $n_1 \le n_2$.\\
\hspace*{1.5em}\textbf{b.} If $n_1<n_2$, lift $\mathcal{U}_1$ to domain
$n_2$ by replication or domain lifting.\\
\hspace*{1.5em}\textbf{c.} Interleave the aligned bundles coordinate-wise to
obtain one bundle $\mathcal{U}_{12}$ of length $n_1+n_2$; if needed, pad it to
the next power of two.\\
\hspace*{1.5em}\textbf{d.} Insert $\mathcal{U}_{12}$ back into the bucket
corresponding to its new domain size.\\
\textbf{3.} Output the final remaining bundle.
\end{quote}

This scheduler is deliberately greedy: by always merging the two smallest
bundles first, it keeps intermediate domains close to the minimum required
size. In the common case where all bundles are eventually padded to a power of
two anyway, this heuristic avoids creating unnecessarily large intermediate
objects.

\subsection{Committed Forest Representation}

We commit the forest using the hyperbox encoding of \cite{cq++24}. Nodes are
indexed so that internal nodes lie in $[N_{\mathrm{int}}]$ and leaves lie in
$\{N_{\mathrm{int}},\ldots,N_{\mathrm{tot}}-1\}$. Trees are ordered first by
boosting round and then by class, so tree $(tK+k)$ is the tree trained for
class $k$ in round $t$.

The committed forest consists of the following objects.
\begin{itemize}
  \item Sparse routing matrices
  $\mathbf{L},\mathbf{R}\in\{0,1\}^{N_{\mathrm{tot}}\times N_{\mathrm{tot}}}$
  encoding left and right children.
  \item Lower and upper hyperbox bounds
  $\mathbf{N}^{\leftarrow},\mathbf{N}^{\rightarrow}
  \in [B+1]^{N_{\mathrm{tot}}\times d}$.
  \item A feature-selection matrix
  $\mathbf{E}\in\{0,1\}^{N_{\mathrm{tot}}\times d}$ whose support records the
  split feature of each internal node.
  \item A threshold tensor
  $\boldsymbol{\Theta}\in\{0,1\}^{N_{\mathrm{tot}}\times d\times B}$ whose support
  records the selected threshold for each internal node.
  \item A tree-index vector
  $\mathbf{tree}\in[TK]^{N_{\mathrm{tot}}}$ and a sparse root matrix
  $\mathbf{Root}\in\{0,1\}^{TK\times N_{\mathrm{tot}}}$.
  \item A leaf-score vector
  $\mathbf{score}\in\mathbb{F}^{N_{\mathrm{tot}}}$.
\end{itemize}

The well-formedness relation includes the standard booleanity, padding, and
single-selection checks, together with the geometric consistency constraints
\begin{align}
\mathbf{Root}\mathbf{N}^{\leftarrow} &= \mathbf{0}, &
\mathbf{Root}\mathbf{N}^{\rightarrow} &= (B+1)\mathbf{1}, \label{eq:root-boxes}\\
\mathbf{Root}\mathbf{tree} &= (0,1,\ldots,TK-1), \label{eq:root-tree}\\
\mathbf{L}\mathbf{tree} &= \mathbf{R}\mathbf{tree}
= \mathbf{tree}_{\mathrm{int}}. \label{eq:tree-consistency}
\end{align}
The remaining hyperbox equations are exactly the same as in \cite{cq++24}: the
selected coordinate is split at the chosen threshold, while every non-selected
coordinate is inherited unchanged by both children.

\subsection{Batched Forest Inference}

Let $\mathbf{X}\in[B]^{b\times d}$ be a batch of query inputs. We write
$\mathbf{X}^{\mathrm{rep}}$ for the row-replicated tensor that repeats
$\mathbf{X}$ once per tree. By \cref{lem:row-replication}, the verifier can
reason about $\mathbf{X}^{\mathrm{rep}}$ through the replicated encoding
obtained by the substitution $X \mapsto X^{TK}$ rather than asking for a
separate commitment.

The prover computes the accepting leaf of every sample in every tree and proves
correctness in two steps.
\begin{enumerate}
  \item It proves a concatenated matrix lookup
  \begin{equation}\label{eq:inference-lookup}
  \mathbf{ind}\ \|\ \mathbf{U}^{\leftarrow}\ \|\ \mathbf{U}^{\rightarrow}
  \ \|\ \mathbf{s}
  \subseteq
  \mathbf{tree}_{\mathrm{ext}}\ \|\ \mathbf{F}^{\leftarrow}\ \|\ 
  \mathbf{F}^{\rightarrow}\ \|\ \mathbf{score},
  \end{equation}
  where each row on the left fetches, for one tree-sample pair, the tree index,
  the lower and upper hyperbox bounds, and the leaf score.
  \item It proves the range relation
  \begin{equation}\label{eq:inference-range}
  \mathbf{U}^{\leftarrow} \leq \mathbf{X}^{\mathrm{rep}} <
  \mathbf{U}^{\rightarrow}.
  \end{equation}
\end{enumerate}

The fetched scores are then summed across trees to obtain the final class-score
vector for each sample. Operationally, this is aggregation along the tree
dimension after the tensor is reshaped into round and class axes.

\subsection{Training Proof}

The training proof is best read as a proof graph rather than as one monolithic
argument. Each round produces a small family of committed tensors, and each
tensor is tied to earlier ones by a short list of recurring checks: lookup,
aggregation, prefix sums, zero-safe divisions, and max-selection constraints.

\paragraph{Round state.}
For each round $t \in [T]$, we maintain:
\begin{itemize}
  \item the cumulative score matrix $\mathbf{S}^{(t)} \in \mathbb{F}^{M\times K}$;
  \item the pseudo-residual matrix $\mathbf{G}^{(t)} \in \mathbb{F}^{M\times K}$;
  \item for every class $k$, the leaf-index vector
  $\boldsymbol{\ell}^{(t,k)} \in [N_{\mathrm{ext}}]^M$ induced by the tree
  trained in round $t$ for class $k$;
  \item the per-node histogram tensors derived from
  $\boldsymbol{\ell}^{(t,k)}$ and $\mathbf{G}^{(t)}[\cdot,k]$;
  \item the selected feature and threshold of every internal node.
\end{itemize}

The proof for one round is organized as follows.

\paragraph{1. Score accumulation and pseudo-residuals.}
For $t=0$, the protocol sets $\mathbf{S}^{(0)}=\mathbf{0}$. For every
$t \geq 1$, let $\mathbf{W}^{(t-1)} \in \mathbb{F}^{N_{\mathrm{ext}}\times K}$
be the matrix of leaf predictions from the previous round. The cumulative
scores satisfy
\begin{equation}\label{eq:score-update}
\mathbf{S}^{(t)}[r,k]
=
\mathbf{S}^{(t-1)}[r,k]
+ \mu \cdot \mathbf{W}^{(t-1)}[\boldsymbol{\ell}^{(t-1,k)}[r],k].
\end{equation}
The right-hand side is proved by a lookup from the leaf-index vector into
$\mathbf{W}^{(t-1)}$, followed by a linear consistency check.

To derive pseudo-residuals, the prover commits auxiliary matrices
$\mathbf{P}^{(t)}, \mathbf{Z}^{(t)}, \mathbf{I}^{(t)}$ satisfying
\begin{align}
\mathbf{P}^{(t)}[r,k] &= \exp(\mathbf{S}^{(t)}[r,k]), \label{eq:softmax-exp}\\
\mathbf{Z}^{(t)}[r] &= \sum_{c=0}^{K-1}\mathbf{P}^{(t)}[r,c], \label{eq:softmax-sum}\\
\mathbf{I}^{(t)}[r] &= \big(\mathbf{Z}^{(t)}[r]\big)^{-1}, \label{eq:softmax-inv}
\end{align}
and then proves
\begin{equation}\label{eq:residual-definition}
\mathbf{G}^{(t)}[r,k]
=
\mathbf{Y}[r,k]-\mathbf{P}^{(t)}[r,k]\mathbf{I}^{(t)}[r].
\end{equation}
Equations \eqref{eq:softmax-exp} and \eqref{eq:softmax-inv} are handled by
lookup-based nonlinear checks or an imported zkML subroutine such as
\cite{zkcnn21,zkdnn24}. Equation \eqref{eq:softmax-sum} is a
matrix-sum check over the class dimension after viewing $\mathbf{P}^{(t)}$ as a
tensor of shape $M\times K$.

\paragraph{2. Leaf-level histograms.}
For each class $k$, define the address tensor
\begin{equation}\label{eq:hist-address}
\mathbf{addr}^{(t,k)}[r,j]
:=
(\boldsymbol{\ell}^{(t,k)}[r]d+j)B + \mathbf{D}[r,j].
\end{equation}
The prover commits a count histogram $\mathbf{C}^{(t,k)}$ and a weighted
histogram $\mathbf{H}^{(t,k)}$ of shape $N_{\mathrm{ext}}\times d\times B$
such that
\begin{align}
\mathbf{C}^{(t,k)}[u,j,b]
&=
\sum_{r=0}^{M-1}
\mathbf{1}\!\left[
\boldsymbol{\ell}^{(t,k)}[r]=u \wedge \mathbf{D}[r,j]=b
\right], \label{eq:leaf-count-hist}\\
\mathbf{H}^{(t,k)}[u,j,b]
&=
\sum_{r=0}^{M-1}
\mathbf{G}^{(t)}[r,k]\,
\mathbf{1}\!\left[
\boldsymbol{\ell}^{(t,k)}[r]=u \wedge \mathbf{D}[r,j]=b
\right]. \label{eq:leaf-weight-hist}
\end{align}
Equations \eqref{eq:leaf-count-hist} and \eqref{eq:leaf-weight-hist} are the
two histogram constraints that drive the rest of the protocol: one for counts
and one for weighted sums.

\paragraph{3. Internal-node propagation.}
Leaf histograms determine internal-node histograms. After flattening the last
two dimensions, the internal tensors satisfy
\begin{align}
\mathbf{C}^{(t,k)}_{\mathrm{int}}
&= \mathbf{L}\mathbf{C}^{(t,k)} + \mathbf{R}\mathbf{C}^{(t,k)},
\label{eq:internal-C}\\
\mathbf{H}^{(t,k)}_{\mathrm{int}}
&= \mathbf{L}\mathbf{H}^{(t,k)} + \mathbf{R}\mathbf{H}^{(t,k)}.
\label{eq:internal-H}
\end{align}
These are linear identities and are absorbed into the batched polynomial
identity used for the round. If the tensors involved live on different subgroup
sizes because of padding, \cref{lem:domain-lifting} lifts them to a common
ambient domain first.

\paragraph{4. Prefix statistics and candidate split scores.}
For each class $k$, the prover next commits prefix tensors
$\mathbf{C}^{(t,k)}_{\leq}, \mathbf{H}^{(t,k)}_{\leq}$ satisfying
\begin{align}
\mathbf{C}^{(t,k)}_{\leq}[u,j,b]
&=
\sum_{q=0}^{b}\mathbf{C}^{(t,k)}[u,j,q], \label{eq:prefix-count}\\
\mathbf{H}^{(t,k)}_{\leq}[u,j,b]
&=
\sum_{q=0}^{b}\mathbf{H}^{(t,k)}[u,j,q]. \label{eq:prefix-weight}
\end{align}
These are prefix-sum checks along the bin axis. The
corresponding totals
\begin{align}
\mathbf{C}^{(t,k)}_{\Sigma}[u,j]
&=
\sum_{q=0}^{B-1}\mathbf{C}^{(t,k)}[u,j,q], &
\mathbf{H}^{(t,k)}_{\Sigma}[u,j]
&=
\sum_{q=0}^{B-1}\mathbf{H}^{(t,k)}[u,j,q]
\end{align}
are ordinary sums over the bin dimension. The right-side statistics are
then derived by subtraction:
\begin{align}
\mathbf{C}^{(t,k)}_{>}[u,j,b]
&=
\mathbf{C}^{(t,k)}_{\Sigma}[u,j]-\mathbf{C}^{(t,k)}_{\leq}[u,j,b], \\
\mathbf{H}^{(t,k)}_{>}[u,j,b]
&=
\mathbf{H}^{(t,k)}_{\Sigma}[u,j]-\mathbf{H}^{(t,k)}_{\leq}[u,j,b].
\end{align}

To avoid undefined divisions, the prover commits a validity mask
$\boldsymbol{\chi}^{(t,k)} \in \{0,1\}^{N_{\mathrm{int}}\times d\times B}$ such that
\[
\boldsymbol{\chi}^{(t,k)}[u,j,b]=1
\iff
\mathbf{C}^{(t,k)}_{\leq}[u,j,b] > 0
\ \wedge\
\mathbf{C}^{(t,k)}_{>}[u,j,b] > 0.
\]
The candidate split-score tensor is then defined by the zero-safe identity
\begin{equation}\label{eq:gain-tensor}
\Gamma^{(t,k)}[u,j,b]
=
\boldsymbol{\chi}^{(t,k)}[u,j,b]
\begin{aligned}[t]
\Bigg(
&\frac{\big(\mathbf{H}^{(t,k)}_{\leq}[u,j,b]\big)^2}
{\mathbf{C}^{(t,k)}_{\leq}[u,j,b] + 1 - \boldsymbol{\chi}^{(t,k)}[u,j,b]} \\
&+
\frac{\big(\mathbf{H}^{(t,k)}_{>}[u,j,b]\big)^2}
{\mathbf{C}^{(t,k)}_{>}[u,j,b] + 1 - \boldsymbol{\chi}^{(t,k)}[u,j,b]}
\Bigg).
\end{aligned}
\end{equation}
When the mask equals one, \eqref{eq:gain-tensor} is exactly the regression-tree
split score. When the mask equals zero, the score is forced to zero and the
division remains well defined.

\paragraph{5. Optimal split selection.}
For each internal node, the prover commits feature-wise maxima
$\mathbf{R}^{(t,k)} \in \mathbb{F}^{N_{\mathrm{int}}\times d}$ and node-wise
maxima $\mathbf{M}^{(t,k)} \in \mathbb{F}^{N_{\mathrm{int}}}$ satisfying
\begin{align}
\mathbf{R}^{(t,k)}[u,j] &= \max_{b \in [B]} \Gamma^{(t,k)}[u,j,b],
\label{eq:feature-max}\\
\mathbf{M}^{(t,k)}[u] &= \max_{j \in [d]} \mathbf{R}^{(t,k)}[u,j].
\label{eq:node-max}
\end{align}
The selected feature matrix $\mathbf{E}$ and threshold tensor $\boldsymbol{\Theta}$
must satisfy the following constraints for every internal node $u$ belonging to
tree $(t,k)$:
\begin{align}
\sum_{j=0}^{d-1}\mathbf{E}[u,j] &= 1,
\qquad
\sum_{b=0}^{B-1}\boldsymbol{\Theta}[u,j,b] = \mathbf{E}[u,j], \label{eq:selection-onehot}\\
\mathbf{E}[u,j] = 1 &\Longrightarrow \mathbf{R}^{(t,k)}[u,j] = \mathbf{M}^{(t,k)}[u],
\label{eq:feature-argmax}\\
\boldsymbol{\Theta}[u,j,b] = 1 &\Longrightarrow \Gamma^{(t,k)}[u,j,b] = \mathbf{R}^{(t,k)}[u,j].
\label{eq:threshold-argmax}
\end{align}
Equations \eqref{eq:feature-max}--\eqref{eq:threshold-argmax} can be discharged
with standard max-selection gadgets over committed tensors. Their role is to
show that the committed hyperboxes are updated using an actual maximizer of the
candidate split-score tensor, rather than an arbitrary split.

\paragraph{6. Leaf values and boosting update.}
The same leaf-index vector also induces per-leaf counts and residual sums:
\begin{align}
\mathbf{n}^{(t,k)}[u]
&=
\sum_{r=0}^{M-1}\mathbf{1}\!\left[\boldsymbol{\ell}^{(t,k)}[r]=u\right],
\label{eq:leaf-count}\\
\mathbf{a}^{(t,k)}[u]
&=
\sum_{r=0}^{M-1}\mathbf{G}^{(t)}[r,k]\,
\mathbf{1}\!\left[\boldsymbol{\ell}^{(t,k)}[r]=u\right]. \label{eq:leaf-sum}
\end{align}
Both are obtained by aggregating per-sample contributions according to the leaf
index. The leaf-value vector
$\mathbf{W}^{(t)}[\cdot,k]$ is then defined by
\begin{equation}\label{eq:leaf-value}
\mathbf{W}^{(t)}[u,k]
=
\frac{\mathbf{a}^{(t,k)}[u]}
{\mathbf{n}^{(t,k)}[u]},
\end{equation}
again with the standard zero-safe masking on padded or unused leaves. Finally,
the next-round score matrix is proved using \eqref{eq:score-update}.

\paragraph{Why this decomposition helps.}
The training proof avoids re-expressing split search as one large arithmetic
circuit. The only nonlinear tensors that are specific to GBDT are
\eqref{eq:gain-tensor} and the max-selection relations derived from it. The
remaining pieces reduce to a small set of reusable checks: lookup,
aggregation, histogram construction, prefix sums, and batched polynomial
identities.

\subsection{Soundness Composition}

The soundness story follows the same decomposition. Any accepting proof for an
incorrect statement must violate at least one of the following:
\begin{itemize}
  \item the soundness of KZG opening proofs;
  \item the soundness of the imported lookup or max-selection subprotocols;
  \item the random-linear-combination soundness of domain lifting and batching;
  \item the random-evaluation soundness of aggregation.
\end{itemize}
For aggregation, the standard check is to compare
\[
\sum_{a=0}^{N-1}\mathbf{h}[a]\rho^a
\quad\text{with}\quad
\sum_{t=0}^{n-1}\mathbf{v}[t]\rho^{\mathbf{ind}[t]}
\]
at a random challenge $\rho$. If the aggregation is incorrect, the difference is
a nonzero polynomial of degree at most $N-1$, so the probability of accidental
acceptance is at most $(N-1)/|\mathbb{F}|$.

Likewise, if a round uses at most $s$ lifted identities under a shared batching
challenge and aggregation range at most $N$, the algebraic contribution to the
round soundness error is
\[
\frac{s-1}{|\mathbb{F}|} + \frac{N-1}{|\mathbb{F}|},
\]
plus the error terms of the imported subprotocols. Domain lifting does not
introduce an additional soundness term beyond the random linear combination in
\eqref{eq:domain-lifted-batch}; it only rewrites smaller-domain claims into the
same ambient divisibility check.
