\section{ZK-GBDT Protocol}\label{sec:zkgbdt-protocol}

This section instantiates the tools of \cref{sec:tools} in the ZK-GBDT
protocol. Appendix~\ref{app:pseudocode} gives pseudocode for the round
structure of quantized GBDT training and proof generation. Here we focus on
the committed objects and relations that the proof must certify.

\subsection{Committed Forest Representation}

We commit the forest using the hyperbox encoding of \cite{cq++24}. Nodes are
indexed from one: internal nodes lie in $\{1,\ldots,N_{\mathrm{int}}\}$ and
leaves lie in $\{N_{\mathrm{int}}+1,\ldots,N_{\mathrm{tot}}\}$. The value $0$
is reserved as the null child pointer for leaf and padding rows. When a node
identifier $u$ labels a committed row, its physical row in the flattened array
is $u-1$. Trees are ordered first by
boosting round and then by class, so tree $(tK+k)$ is the tree trained for
class $k$ in round $t$.

The committed forest consists of the following objects.
\begin{itemize}
  \item Left and right child index vectors
  $\mathbf{lc},\mathbf{rc}\in\{0,1,\ldots,N_{\mathrm{tot}}\}^{N_{\mathrm{tot}}}$
  encoding the one-based node identifiers of the left and right child of each
  internal node.
  \item Lower and upper hyperbox bounds
  $\mathbf{N}^{\leftarrow},\mathbf{N}^{\rightarrow}
  \in [B+1]^{N_{\mathrm{tot}}\times d}$.
  \item A feature-selection matrix
  $\mathbf{E}\in\{0,1\}^{N_{\mathrm{tot}}\times d}$ whose support records the
  split feature of each internal node.
  \item A threshold-index vector
  $\boldsymbol{\Theta}\in[B]^{N_{\mathrm{tot}}}$ whose internal-node entries
  record the selected threshold bin.
  \item A tree-index vector
  $\mathbf{tree}\in[TK]^{N_{\mathrm{tot}}}$ and a sparse root matrix
  $\mathbf{Root}\in\{0,1\}^{TK\times N_{\mathrm{tot}}}$.
  \item A leaf-score vector
  $\mathbf{score}\in\mathbb{F}^{N_{\mathrm{tot}}}$.
\end{itemize}
Whenever a matrix form is convenient, we write
$\mathbf{L},\mathbf{R}\in\{0,1\}^{N_{\mathrm{tot}}\times N_{\mathrm{tot}}}$
for the virtual sparse routing operators induced by
$\mathbf{lc},\mathbf{rc}$, namely
$\mathbf{L}[u,v]=1 \iff \mathbf{lc}[u]=v$ and
$\mathbf{R}[u,v]=1 \iff \mathbf{rc}[u]=v$ for nonzero node identifiers
$u,v\in\{1,\ldots,N_{\mathrm{tot}}\}$.

\begin{proposition}[Forest well-formedness]
\label{prop:forest-well-formedness}
The committed forest is well formed if the following hold:
\begin{enumerate}
  \item every active internal entry of
  $\mathbf{lc},\mathbf{rc}$ points to one child row,
  leaf and padding entries of these vectors are zero, and every active
  internal row of $\mathbf{E}$ is one-hot while leaf and padding rows are
  zero;
  \item $\boldsymbol{\Theta}$ is zero on leaf and padding rows, and
  $\mathbf{score}$ is zero on internal and padding rows;
  \item the child rows selected by
  $\mathbf{lc},\mathbf{rc}$ belong to the same tree
  as their parent, and the root selector $\mathbf{Root}$ chooses exactly one
  root row per tree;
  \item the selected coordinate updates exactly one box bound at the chosen
  threshold, while every non-selected coordinate is inherited unchanged by both
  children;
  \item root rows carry the global box, and inactive padding rows are zeroed.
\end{enumerate}
In matrix form, the root and tree-consistency constraints are
\begin{align}
\mathbf{Root}\mathbf{N}^{\leftarrow} &= \mathbf{0}_{TK\times d}, &
\mathbf{Root}\mathbf{N}^{\rightarrow} &= (B+1)\mathbf{1}_{TK\times d},
\label{eq:root-boxes}\\
\mathbf{Root}\mathbf{tree} &= (0,1,\ldots,TK-1)^\top, \label{eq:root-tree}\\
\mathbf{L}\mathbf{tree} &= \mathbf{R}\mathbf{tree}
= \mathbf{tree}_{\mathrm{int}}. \label{eq:tree-consistency}
\end{align}
Here $\mathbf{tree}_{\mathrm{int}}\in[TK]^{N_{\mathrm{tot}}}$ denotes the
tree-index vector on internal rows, extended by zeros on leaf and padding rows.
\end{proposition}

\subsection{Batched Forest Inference}

Let $\mathbf{X}\in[B]^{b\times d}$ be a batch of query inputs. Depending on the
surrounding application, $\mathbf{X}$ may be public or may itself be a
committed witness object. We write
$\mathbf{X}^{\mathrm{rep}}$ for the row-replicated tensor that repeats
$\mathbf{X}$ once per tree. By \cref{lem:row-replication}, the verifier can
reason about $\mathbf{X}^{\mathrm{rep}}$ through the replicated encoding
obtained by the substitution $X \mapsto X^{TK}$ rather than asking for a
separate commitment.

The inference proof has two stages: row fetch and box containment. For every
tree-sample pair, the prover identifies the accepting leaf and then proves that
this leaf is the correct one.
\begin{enumerate}
  \item It proves a concatenated matrix lookup
  \begin{equation}\label{eq:inference-lookup}
  \mathbf{ind}\ \|\ \mathbf{U}^{\leftarrow}\ \|\ \mathbf{U}^{\rightarrow}
  \ \|\ \mathbf{s}
  \subseteq
  \mathbf{tree}_{\mathrm{ext}}\ \|\ \mathbf{F}^{\leftarrow}\ \|\ 
  \mathbf{F}^{\rightarrow}\ \|\ \mathbf{score},
  \end{equation}
  where
  $\mathbf{ind}\in[TK]^{bTK}$,
  $\mathbf{U}^{\leftarrow},\mathbf{U}^{\rightarrow}\in[B+1]^{bTK\times d}$,
  and $\mathbf{s}\in\mathbb{F}^{bTK}$ are the fetched tree indices, lower and
  upper boxes, and leaf scores for the $bTK$ tree-sample pairs; and
  $\mathbf{tree}_{\mathrm{ext}}, \mathbf{F}^{\leftarrow},
  \mathbf{F}^{\rightarrow}$ denote the leaf-row restrictions of
  $\mathbf{tree}, \mathbf{N}^{\leftarrow}, \mathbf{N}^{\rightarrow}$.
  \item It proves the range relation
  \begin{equation}\label{eq:inference-range}
  \mathbf{U}^{\leftarrow} \leq \mathbf{X}^{\mathrm{rep}} <
  \mathbf{U}^{\rightarrow}.
  \end{equation}
\end{enumerate}

\begin{proposition}[Batched forest inference]
\label{prop:batched-forest-inference}
Let $\mathbf{tree}_{\mathrm{ext}}$,
$\mathbf{F}^{\leftarrow}$, and
$\mathbf{F}^{\rightarrow}$ denote the restrictions of
$\mathbf{tree}$, $\mathbf{N}^{\leftarrow}$, and
$\mathbf{N}^{\rightarrow}$ to leaf rows only. If
\eqref{eq:inference-lookup} and \eqref{eq:inference-range} hold and the forest
is well formed, then every tree-sample pair fetches exactly one leaf row from
the correct tree, and that row is the unique accepting leaf for the replicated
query.
\end{proposition}

This same statement is reused later for the training witness
$\boldsymbol{\ell}^{(t,k)}$: proving leaf assignments is nothing more than
running the same fetch-and-containment argument on the training dataset instead
of on an external inference batch.

The fetched leaf scores are then summed across trees to obtain the final
class-score vector of each sample. Operationally, this is aggregation along the
tree axis after reshaping into round and class axes.

\subsection{Round-Level Training Proof}

The training proof is best read as a round-level dependency graph rather than a
single monolithic argument. Each round introduces a small family of committed
tensors, and each one is linked to earlier objects through a short list of
recurring checks: lookup, aggregation, prefix sums, zero-safe divisions, and
max-selection constraints.

\paragraph{Round state.}
At the commitment layer, the boosting state is represented by global tensors
\[
\mathbf{S}\in\mathbb{F}^{T\times K\times M},\qquad
\mathbf{G}\in\mathbb{F}^{T\times K\times M},\qquad
\mathbf{P}\in\mathbb{F}^{T\times K\times M},
\]
\[
\mathbf{Z}\in\mathbb{F}^{T\times M},\qquad
\mathbf{I}\in\mathbb{F}^{T\times M},
\]
all flattened in row-major order before commitment. The score and residual
tensors use the same round-class-sample order as the training overview:
$\mathbf{G}[t,k,r]$ is the target used by the class-$k$
tree in round $t$, and $\mathbf{S}[t,k,r]$ is the cumulative score after that
round. We write the virtual initial state as $\mathbf{S}[-1,k,r]=0$. The point
of this arrangement is that the number of PCS instances and the opening pattern
do not depend on $T$, although the witness size and prover work still do.

For each round $t \in [T]$ we maintain:
\begin{itemize}
  \item the cumulative score slice
  $\mathbf{S}[t,\cdot,\cdot] \in \mathbb{F}^{K\times M}$;
  \item the pseudo-residual slice
  $\mathbf{G}[t,\cdot,\cdot] \in \mathbb{F}^{K\times M}$;
  \item for every class $k$, the leaf-index vector
  $\boldsymbol{\ell}^{(t,k)} \in
  \{N_{\mathrm{int}}+1,\ldots,N_{\mathrm{tot}}\}^{M}$ induced by the tree
  trained in round $t$ for class $k$;
  \item leaf-supported and all-node histogram tensors derived from
  $\boldsymbol{\ell}^{(t,k)}$ and $\mathbf{G}[t,k,\cdot]$;
  \item the selected feature-threshold pair of every internal node.
\end{itemize}
All of these round-indexed views are slices or derived slices of globally
committed objects. In particular, the score-update relations of all rounds are
batched into one shared boosting-state proof rather than proved in separate PCS
instances.

\paragraph{Imported softmax interface.}
We treat softmax as an imported subprotocol interface. Concretely, the main
protocol assumes a proof system $\Pi_{\mathrm{softmax}}$ which, given committed
score, exponential, denominator, and inverse tensors, proves the relations
\eqref{eq:softmax-exp}--\eqref{eq:softmax-inv} with its own
completeness, soundness, and zero-knowledge guarantees. The security statement
of ZK-GBDT is parameterized by this imported interface.

Likewise, we treat lookup, aggregation, and max-selection as named imported
interfaces $\Pi_{\mathrm{lookup}}$, $\Pi_{\mathrm{agg}}$, and
$\Pi_{\mathrm{max}}$. The new contribution of this paper is therefore not these
gadgets in isolation, but the way Sections~4--5 compose them with
domain-lifting batching, interleaving batching, and the GBDT-specific tensor
relations.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[
    node distance=0.95cm and 0.55cm,
    box/.style={draw, rounded corners, align=center, minimum height=8mm, inner sep=4pt, fill=blue!3},
    arr/.style={-{Latex[length=2mm]}, thick}
  ]
    \node[box] (scores) {$\mathbf{S}[t-1,\cdot,\cdot]$};
    \node[box, right=of scores] (resid) {$\mathbf{G}[t,\cdot,\cdot]$};
    \node[box, right=of resid] (leaf) {$\boldsymbol{\ell}^{(t,k)}$};
    \node[box, below=of leaf] (hist) {$\mathbf{C}^{\mathrm{leaf},(t,k)},\ \mathbf{H}^{\mathrm{leaf},(t,k)}$};
    \node[box, left=of hist] (allnode) {$\mathbf{C}^{(t,k)},\ \mathbf{H}^{(t,k)}$};
    \node[box, left=of allnode] (gain) {$\Gamma^{(t,k)}$};
    \node[box, below=of gain] (split) {$\mathbf{E},\ \boldsymbol{\Theta}$};
    \node[box, right=of split] (leafvalue) {$\widetilde{\mathbf{W}}^{(t)}$};
    \node[box, right=of leafvalue] (nextscores) {$\mathbf{S}[t,\cdot,\cdot]$};
    \draw[arr] (scores) -- node[above, font=\small] {softmax} (resid);
    \draw[arr] (resid) -- node[above, font=\small] {batched inference} (leaf);
    \draw[arr] (leaf) -- node[right, font=\small] {histogram agg.} (hist);
    \draw[arr] (hist) -- node[above, font=\small] {child-index pull-up} (allnode);
    \draw[arr] (allnode) -- node[above, font=\small] {prefix / zero-safe score} (gain);
    \draw[arr] (gain) -- node[right, font=\small] {argmax bind} (split);
    \draw[arr] (hist) -- node[right, font=\small] {leaf aggregation} (leafvalue);
    \draw[arr] (leafvalue) -- node[above, font=\small] {boosting update} (nextscores);
  \end{tikzpicture}
  \caption{Round-level proof flow for one boosting round. The same leaf
  assignments support both split certification and the computation of the
  current round's leaf values.}
  \label{fig:round-proof-flow}
\end{figure}

\noindent\Cref{fig:round-proof-flow} shows the dependency structure of one
round. The same leaf assignments drive two downstream tasks: they generate the
histograms used for split certification, and they determine the per-leaf
aggregates that define the current round's leaf values.

\begin{proposition}[Boosting-state and pseudo-residual correctness]
\label{prop:boosting-state}
The protocol uses the virtual initial state $\mathbf{S}[-1,k,r]=0$ and sets
$\mathbf{G}[0,k,r]=\mathbf{Y}[k,r]$. For every $t\in[T]$, let
$\widetilde{\mathbf{W}}^{(t)} \in \mathbb{F}^{N_{\mathrm{tot}}\times K}$
denote the padded leaf-value matrix from round $t$, with zeros on internal
rows. The cumulative scores satisfy
\begin{equation}\label{eq:score-update}
\mathbf{S}[t,k,r]
=
\mathbf{S}[t-1,k,r]
+ \mu \cdot \widetilde{\mathbf{W}}^{(t)}[\boldsymbol{\ell}^{(t,k)}[r],k].
\end{equation}
The right-hand side is proved by a lookup from the leaf-index vector into
$\widetilde{\mathbf{W}}^{(t)}$, followed by a linear consistency check.

For $t<T-1$, the next pseudo-residual slice is derived from the score slice
after round $t$. The prover commits auxiliary tensors satisfying
\begin{align}
\mathbf{P}[t,k,r] &= \exp(\mathbf{S}[t,k,r]), \label{eq:softmax-exp}\\
\mathbf{Z}[t,r] &= \sum_{c=0}^{K-1}\mathbf{P}[t,c,r], \label{eq:softmax-sum}\\
\mathbf{I}[t,r] &= \big(\mathbf{Z}[t,r]\big)^{-1}, \label{eq:softmax-inv}
\end{align}
and then proves
\begin{equation}\label{eq:residual-definition}
\mathbf{G}[t+1,k,r]
=
\mathbf{Y}[k,r]-\mathbf{P}[t,k,r]\mathbf{I}[t,r].
\end{equation}
Equation \eqref{eq:softmax-sum} is checked directly as a linear relation.
Equations \eqref{eq:softmax-exp} and \eqref{eq:softmax-inv} are delegated to
the imported interface $\Pi_{\mathrm{softmax}}$. Hence the protocol certifies
that $(\mathbf{S},\mathbf{G})$ is consistent with the quantized
trainer semantics of Section~2.
\end{proposition}

\begin{proposition}[Leaf assignment correctness]
\label{prop:leaf-assignment}
For every round-class pair $(t,k)$, the vector
$\boldsymbol{\ell}^{(t,k)}\in
\{N_{\mathrm{int}}+1,\ldots,N_{\mathrm{tot}}\}^{M}$ is proved by applying
\cref{prop:batched-forest-inference} to the training dataset $\mathbf{D}$ and
the tree $(t,k)$. Consequently, $\boldsymbol{\ell}^{(t,k)}[r]$ is the unique
accepting leaf for sample $r$ in that tree.
\end{proposition}

\begin{proposition}[Leaf histogram correctness]
\label{prop:leaf-hist}
For each class $k$, define the address tensor
\begin{equation}\label{eq:hist-address}
\mathbf{addr}^{(t,k)}[r,j]
:=
\bigl((\boldsymbol{\ell}^{(t,k)}[r]-1)d+j\bigr)B + \mathbf{D}[r,j].
\end{equation}
The prover first commits leaf-supported tensors
$\mathbf{C}^{\mathrm{leaf},(t,k)}$ and
$\mathbf{H}^{\mathrm{leaf},(t,k)}$ of shape
$N_{\mathrm{tot}}\times d\times B$. They are supported only on leaf rows and
satisfy
\begin{align}
\mathbf{C}^{\mathrm{leaf},(t,k)}[u,j,b]
&=
\sum_{r=0}^{M-1}
\mathbf{1}\!\left[
\boldsymbol{\ell}^{(t,k)}[r]=u \wedge \mathbf{D}[r,j]=b
\right], \label{eq:leaf-count-hist}\\
\mathbf{H}^{\mathrm{leaf},(t,k)}[u,j,b]
&=
\sum_{r=0}^{M-1}
\mathbf{G}[t,k,r]\,
\mathbf{1}\!\left[
\boldsymbol{\ell}^{(t,k)}[r]=u \wedge \mathbf{D}[r,j]=b
\right]. \label{eq:leaf-weight-hist}
\end{align}
These are the two basic histogram relations used by the rest of the protocol:
one for counts and one for weighted sums. Internal-node rows of these
leaf-supported tensors are constrained to zero. Equivalently, if
$\Flat(\mathbf{C}^{\mathrm{leaf},(t,k)})$ and
$\Flat(\mathbf{H}^{\mathrm{leaf},(t,k)})$ are viewed as length
$N_{\mathrm{tot}}dB$ vectors, then the protocol proves by aggregation that
placing the value $1$ at each address $\mathbf{addr}^{(t,k)}[r,j]$ produces
$\Flat(\mathbf{C}^{\mathrm{leaf},(t,k)})$. Placing the value
$\mathbf{G}[t,k,r]$ at those same addresses produces
$\Flat(\mathbf{H}^{\mathrm{leaf},(t,k)})$.
\end{proposition}

\begin{proposition}[Internal histogram propagation]
\label{prop:internal-propagation}
Leaf-supported histograms determine the full all-node histograms
$\mathbf{C}^{(t,k)}$ and $\mathbf{H}^{(t,k)}$. After flattening the last two
dimensions, they satisfy
\begin{align}
\Flat\!\bigl(\mathbf{C}^{(t,k)}\bigr)
&=
\Flat\!\bigl(\mathbf{C}^{\mathrm{leaf},(t,k)}\bigr)
+ \mathbf{L}\Flat\!\bigl(\mathbf{C}^{(t,k)}\bigr)
+ \mathbf{R}\Flat\!\bigl(\mathbf{C}^{(t,k)}\bigr),
\label{eq:internal-C}\\
\Flat\!\bigl(\mathbf{H}^{(t,k)}\bigr)
&=
\Flat\!\bigl(\mathbf{H}^{\mathrm{leaf},(t,k)}\bigr)
+ \mathbf{L}\Flat\!\bigl(\mathbf{H}^{(t,k)}\bigr)
+ \mathbf{R}\Flat\!\bigl(\mathbf{H}^{(t,k)}\bigr).
\label{eq:internal-H}
\end{align}
These are linear identities, so they are absorbed into the batched polynomial
identity used for the round. If padding places the participating tensors on
different subgroup sizes, \cref{thm:domain-lifting-batching} lifts them to a common
ambient domain first. Because nodes are topologically ordered, the operator
induced by $\mathbf{L}+\mathbf{R}$ is nilpotent on each tree, so these
equations have a unique intended solution: upward accumulation from leaves to
ancestors. The histograms are also masked to zero outside the support of the
tree $(t,k)$.
\end{proposition}

\begin{proposition}[Prefix, total, and right-side statistics]
\label{prop:prefix-total}
Let $\mathsf{Pref}_B$ denote the public linear prefix operator along the bin
axis. For each class $k$, the prover next commits prefix tensors
$\mathbf{C}^{(t,k)}_{\leq}, \mathbf{H}^{(t,k)}_{\leq}
\in\mathbb{F}^{N_{\mathrm{tot}}\times d\times B}$ satisfying
\begin{align}
\mathbf{C}^{(t,k)}_{\leq}[u,j,b]
&=
\sum_{q=0}^{b}\mathbf{C}^{(t,k)}[u,j,q], \label{eq:prefix-count}\\
\mathbf{H}^{(t,k)}_{\leq}[u,j,b]
&=
\sum_{q=0}^{b}\mathbf{H}^{(t,k)}[u,j,q]. \label{eq:prefix-weight}
\end{align}
The corresponding totals
\[
\mathbf{C}^{(t,k)}_{\Sigma},\mathbf{H}^{(t,k)}_{\Sigma}
\in\mathbb{F}^{N_{\mathrm{tot}}\times d}
\]
are
\begin{align}
\mathbf{C}^{(t,k)}_{\Sigma}[u,j]
&=
\sum_{q=0}^{B-1}\mathbf{C}^{(t,k)}[u,j,q], &
\mathbf{H}^{(t,k)}_{\Sigma}[u,j]
&=
\sum_{q=0}^{B-1}\mathbf{H}^{(t,k)}[u,j,q]
\end{align}
ordinary sums over the bin dimension. The right-side statistics
\[
\mathbf{C}^{(t,k)}_{>},\mathbf{H}^{(t,k)}_{>}
\in\mathbb{F}^{N_{\mathrm{tot}}\times d\times B}
\]
are then derived by subtraction:
\begin{align}
\mathbf{C}^{(t,k)}_{>}[u,j,b]
&=
\mathbf{C}^{(t,k)}_{\Sigma}[u,j]-\mathbf{C}^{(t,k)}_{\leq}[u,j,b], \\
\mathbf{H}^{(t,k)}_{>}[u,j,b]
&=
\mathbf{H}^{(t,k)}_{\Sigma}[u,j]-\mathbf{H}^{(t,k)}_{\leq}[u,j,b].
\end{align}
All of these are linear identities, so they are eligible for
domain-lifting batching together with the other tensor-linear relations of the
round.
\end{proposition}

\begin{proposition}[Candidate split-score correctness]
\label{prop:candidate-score}
To avoid undefined divisions, the prover commits a validity mask
$\boldsymbol{\chi}^{(t,k)} \in \{0,1\}^{N_{\mathrm{int}}\times d\times B}$ such that
\[
\boldsymbol{\chi}^{(t,k)}[u,j,b]=1
\iff
\mathbf{C}^{(t,k)}_{\leq}[u,j,b] > 0
\ \wedge\
\mathbf{C}^{(t,k)}_{>}[u,j,b] > 0.
\]
The candidate split-score tensor
$\Gamma^{(t,k)}\in\mathbb{F}^{N_{\mathrm{int}}\times d\times B}$ is then
represented by a zero-safe quotient relation. Concretely, the prover supplies
auxiliary quotient witnesses
$\mathbf{Q}^{\leftarrow,(t,k)},\mathbf{Q}^{\rightarrow,(t,k)}$ and enforces
\begin{align}
\bigl(\mathbf{C}^{(t,k)}_{\leq}+1-\boldsymbol{\chi}^{(t,k)}\bigr)
\odot \mathbf{Q}^{\leftarrow,(t,k)}
&=
\bigl(\mathbf{H}^{(t,k)}_{\leq}\bigr)^{\odot 2}, \\
\bigl(\mathbf{C}^{(t,k)}_{>}+1-\boldsymbol{\chi}^{(t,k)}\bigr)
\odot \mathbf{Q}^{\rightarrow,(t,k)}
&=
\bigl(\mathbf{H}^{(t,k)}_{>}\bigr)^{\odot 2},
\end{align}
followed by
\begin{equation}\label{eq:gain-tensor}
\Gamma^{(t,k)}
=
\boldsymbol{\chi}^{(t,k)} \odot
\bigl(\mathbf{Q}^{\leftarrow,(t,k)}+\mathbf{Q}^{\rightarrow,(t,k)}\bigr).
\end{equation}
When the mask equals one, \eqref{eq:gain-tensor} is exactly the objective in
\eqref{eq:split-objective}. When the mask equals zero, the score is forced to
zero so that the division remains well defined.
\end{proposition}

\begin{proposition}[Split optimality]
\label{prop:split-optimality}
For each internal node, the prover commits feature-wise maxima
$\mathbf{R}^{(t,k)} \in \mathbb{F}^{N_{\mathrm{int}}\times d}$ and node-wise
maxima $\mathbf{M}^{(t,k)} \in \mathbb{F}^{N_{\mathrm{int}}}$ satisfying
\begin{align}
\mathbf{R}^{(t,k)}[u,j] &= \max_{b \in [B]} \Gamma^{(t,k)}[u,j,b],
\label{eq:feature-max}\\
\mathbf{M}^{(t,k)}[u] &= \max_{j \in [d]} \mathbf{R}^{(t,k)}[u,j].
\label{eq:node-max}
\end{align}
The selected feature matrix $\mathbf{E}$ and threshold vector
$\boldsymbol{\Theta}$ must satisfy the following constraints for every internal
node $u$ belonging to tree $(t,k)$:
\begin{align}
\sum_{j=0}^{d-1}\mathbf{E}[u,j] &= 1, \label{eq:selection-onehot}\\
\mathbf{E}[u,j] = 1 &\Longrightarrow \mathbf{R}^{(t,k)}[u,j] = \mathbf{M}^{(t,k)}[u],
\label{eq:feature-argmax}\\
\mathbf{E}[u,j] = 1 &\Longrightarrow
\Gamma^{(t,k)}[u,j,\boldsymbol{\Theta}[u]] = \mathbf{R}^{(t,k)}[u,j].
\label{eq:threshold-argmax}
\end{align}
Equations \eqref{eq:feature-max}--\eqref{eq:threshold-argmax} are delegated to
the imported interface $\Pi_{\mathrm{max}}$ over committed tensors. The first
relation identifies the maximizing feature, and the second fixes the maximizing
threshold bin on that feature. Under the trainer semantics fixed in
Section~2, this proves that $(\mathbf{E},\boldsymbol{\Theta})$ is an arbitrary
maximizing valid split; ties are not further canonicalized.
\end{proposition}

\begin{proposition}[Leaf values and round closure]
\label{prop:leaf-values}
The same leaf-index vector also induces per-leaf counts and residual sums:
\begin{align}
\widetilde{\mathbf{n}}^{(t,k)}[u]
&=
\sum_{r=0}^{M-1}\mathbf{1}\!\left[\boldsymbol{\ell}^{(t,k)}[r]=u\right],
\label{eq:leaf-count}\\
\widetilde{\mathbf{a}}^{(t,k)}[u]
&=
\sum_{r=0}^{M-1}\mathbf{G}[t,k,r]\,
\mathbf{1}\!\left[\boldsymbol{\ell}^{(t,k)}[r]=u\right]. \label{eq:leaf-sum}
\end{align}
Both are obtained by aggregating per-sample contributions according to the
global leaf index. Here
$\widetilde{\mathbf{n}}^{(t,k)},\widetilde{\mathbf{a}}^{(t,k)}
\in\mathbb{F}^{N_{\mathrm{tot}}}$.
The padded leaf-value matrix
$\widetilde{\mathbf{W}}^{(t)}\in\mathbb{F}^{N_{\mathrm{tot}}\times K}$ is then
defined by a second zero-safe division relation; equivalently, the protocol
enforces
\begin{equation}\label{eq:leaf-value}
\widetilde{\mathbf{n}}^{(t,k)}[u]\cdot \widetilde{\mathbf{W}}^{(t)}[u,k]
=
\widetilde{\mathbf{a}}^{(t,k)}[u]
\end{equation}
on active leaves, together with zero masks on padded or unused leaves. The
compact leaf-only matrix
$\mathbf{W}^{(t)}\in\mathbb{F}^{(N_{\mathrm{tot}}-N_{\mathrm{int}})\times K}$
is obtained by restricting
$\widetilde{\mathbf{W}}^{(t)}$ to leaf rows. Finally, the next-round score
matrix is proved using \eqref{eq:score-update}.
\end{proposition}

\begin{theorem}[Round-level correctness]
\label{thm:round-correctness}
Fix one round $t\in[T]$. If
\cref{prop:boosting-state,prop:leaf-assignment,prop:leaf-hist,prop:internal-propagation,prop:prefix-total,prop:candidate-score,prop:split-optimality,prop:leaf-values}
all hold for every class $k\in[K]$, then the committed round-$t$ objects encode
one valid step of the quantized trainer semantics from
the previous score state to $\mathbf{S}[t,\cdot,\cdot]$.
\end{theorem}

\begin{proof}
The cited propositions certify, in order, the correctness of the round input,
the unique routing of every sample to one leaf, the induced histogram
statistics, the reconstruction of the masked split objective, the optimality of
the committed split metadata, and the computation of the leaf values used in
the next score update. Together these are exactly the conditions defining one
valid round of the trainer semantics from Section~2.
\end{proof}

\subsection{Soundness Composition}

For aggregation, the imported interface $\Pi_{\mathrm{agg}}$ checks
\[
\sum_{a=0}^{N-1}\mathbf{h}[a]\rho^a
\quad\text{with}\quad
\sum_{t=0}^{n-1}\mathbf{v}[t]\rho^{\mathbf{ind}[t]}
\]
at a random challenge $\rho$. If the aggregation is incorrect, the difference is
a nonzero polynomial of degree at most $N-1$, so the probability of accidental
acceptance is at most $(N-1)/|\mathbb{F}|$.

Likewise, if a round uses at most $s$ lifted identities under a shared batching
challenge and aggregation range at most $N$, the algebraic contribution to the
round soundness error is
\[
\frac{s-1}{|\mathbb{F}|} + \frac{N-1}{|\mathbb{F}|},
\]
plus the error terms of the imported subprotocols. Domain lifting introduces no
additional soundness term beyond the random linear combination in
\eqref{eq:domain-lifting-batching}; it only rewrites smaller-domain claims into the
same ambient divisibility check.

\begin{theorem}[Protocol soundness]
\label{thm:protocol-soundness}
Assume the KZG opening protocol is sound, the imported lookup, softmax, and
max-selection subprotocols are sound, and all verifier challenges used for
domain-lifting batches, interleaving reductions, and aggregation checks are
sampled independently. Then any accepting proof for a false statement implies a
failure in at least one of the following components:
\begin{enumerate}
  \item KZG binding or opening soundness;
  \item one of the imported subprotocols;
  \item a domain-lifting random linear combination;
  \item an interleaving-batched local check;
  \item an aggregation check.
\end{enumerate}
Consequently, if the protocol uses at most
$N_{\mathrm{agg}}$ aggregation challenges with ranges bounded by
$R_1,\ldots,R_{N_{\mathrm{agg}}}$, at most
$N_{\mathrm{dl}}$ domain-lifting batches with arities
$s_1,\ldots,s_{N_{\mathrm{dl}}}$, and imported subprotocols with total failure
probability $\epsilon_{\mathrm{imp}}$, then the total soundness error is at
most
\[
\epsilon_{\mathrm{KZG}}
\;+\;
\epsilon_{\mathrm{imp}}
\;+\;
\sum_{i=1}^{N_{\mathrm{dl}}}\frac{s_i-1}{|\mathbb{F}|}
\;+\;
\sum_{j=1}^{N_{\mathrm{agg}}}\frac{R_j-1}{|\mathbb{F}|}
\;+\;
\epsilon_{\mathrm{int}},
\]
where $\epsilon_{\mathrm{KZG}}$ is the KZG failure probability and
$\epsilon_{\mathrm{int}}$ is the total soundness error contributed by the
interleaving-batched local gadgets.
We again emphasize that this is a soundness statement, not a separate
argument-of-knowledge claim.
\end{theorem}

\begin{proof}
Combine \cref{thm:domain-lifting-batching} with the standard aggregation bound
above and a union bound over all imported subprotocol invocations and KZG
openings. Independence of the verifier challenges justifies summing the
individual failure probabilities. The interleaving step contributes no extra
algebraic error beyond the soundness error of the local gadgets themselves; it
only rewrites several local instances into one larger virtual instance.
\end{proof}
