\section{Evaluation}\label{sec:evaluation}

\subsection{Evaluation Setup}\label{subsec:eval-setup}

\paragraph{Hardware.} We implement \textsc{Terrae} in C++ and evaluate
it on Ubuntu 24.04.4 LTS (Linux kernel 6.17.0-14-generic, x86\_64).                                                          
The evaluation machine is equipped with two Intel Xeon Gold 6326 CPUs at                                                     
2.90 GHz, providing 32 physical cores and 64 logical cores in total, and                                                     
188 GB of RAM.     

\paragraph{Baseline.}
There are two candidates for the baseline: (1) Sparrow~\cite{sparrow24}, and (2) ZKBoost~\cite{zkboost26}. The first is 
designed for random forests, and the second is designed for XGBoost,
a specific implementation of GBDT.
The two systems share the same design philosophy for
committing forests and proving training and inference.
We adopt ZKBoost as our baseline, as it is more closely aligned with our design and can be easily adapted to support GBDT.
To ensure a fair comparison, we must keep the core design of ZKBoost,
which includes
\begin{enumerate}
    \item The approach to committing forests.
        In ZKBoost, each tree is padded with
        dummy nodes to the same depth
        and to a full binary tree.
        This makes the proof of routing easier,
        since the structure of the forest is fixed and public.
    \item The approach to proving inference.
        In ZKBoost, the prover commits the paths
        the samples take
        from the roots to the leaves,
        validates that the paths are consistent with the committed forest,
        and then proves the correctness of the leaf values.
        This follows the style
        of \cite{zkdt20} for decision tree inference.
    \item The approach to proving training.
        In ZKBoost, the correctness of training is proved
        in a way similar to \textsc{Terrae},
        except that it relies on generic ZKPs for arithmetic circuits, which should be less efficient than our custom-designed proof system.
\end{enumerate}

Our implementation of ZKBoost strictly follows the above design choice of \cite{zkboost26}.
    We let our implementation of ZKBoost
    share the same cryptographic and 
    algebraic primitives as \textsc{Terrae},
    to ensure a fair comparison of the proof systems.

\paragraph{Implementation.} 
We implement \textsc{Terrae} and ZKBoost in C++.
  Cryptographic 
  operations are based on the mcl library~\cite{mcllib}, and the proof system is instantiated over the BLS12-381 pairing-friendly elliptic curve 
  with KZG polynomial commitments. This  
  instantiation targets roughly 128-bit security, following the standard security level of BLS12-381.
  We use Fiat-Shamir transformation to make 
  the proof systems non-interactive.

Our code is available at \url{https://anonymous.4open.science/r/ZK-GBDT-F601}.


\subsection{Asymptotic Comparison with Prior Works}\label{subsec:eval-asymptotic}

\Cref{tab:complexity-comparison} summarizes the asymptotic prover costs of
\textsc{Terrae} and the baseline. 

\input{tables/complexity_comparison}

One significant phenomenon is that the prover cost of training certification in \textsc{Terrae} is asymptotically smaller than the baseline by a factor of $h$.
This is due to the fact that \textsc{Terrae} avoids
committing the entire root-to-leaf path for proving inference, and thus avoids the need to prove the consistency of the path with the committed forest, which is a major cost in the baseline.

\subsection{Empirical Scaling in Three Dimensions}\label{subsec:eval-three-dim}

We next study how proving time scales with three key workload parameters: the
number of iterations $T$, 
the maximum tree depth $h$, 
and the size of training dataset $M$.
We start with a basic case
$(T,h,M)=(1,1,512)$ and scale each parameter while keeping the others fixed.
For proof of training, we set the batch size $b=M$.
\Cref{tab:three-dim-scaling} reports prover time and peak prover memory for
both training and inference.

\input{tables/three_dim_scaling}

\paragraph{Scaling with the number of trees.}
As expected, prover time in both systems grows roughly linearly with $T$, since
each additional tree introduces another proof instance of essentially the same
shape.  Across the entire sweep, \textsc{Terrae} consistently outperforms
ZKBoost by a stable margin: $4.90\times$--$5.23\times$ for training and
$5.52\times$--$5.82\times$ for inference.  From $T=1$ to $T=4$,
\textsc{Terrae}'s proving time increases from $58.79$\,s to $239.96$\,s for
training and from $15.76$\,s to $69.93$\,s for inference, which is close to
linear scaling.  Peak memory grows with $T$ as well, but \textsc{Terrae}
remains consistently lighter: at $T=4$, it uses $16.46$\,GB for training and
$4.83$\,GB for inference, compared with $39.03$\,GB and $16.36$\,GB for
ZKBoost.  This suggests that the improvement of \textsc{Terrae} is not tied to
a narrow parameter regime, but persists as the ensemble grows.

\paragraph{Scaling with tree depth.}
Depth is the dimension where the difference between the two approaches becomes
most pronounced.  On inference, \textsc{Terrae} is nearly insensitive to depth
in this range: its proving time changes only from $15.61$\,s at $h=1$ to
$17.41$\,s at $h=4$.  In contrast, ZKBoost increases from $86.24$\,s to
$413.98$\,s, leading to speedups of $5.53\times$--$23.78\times$.  A similar,
though less uniform, trend appears in training, where \textsc{Terrae} achieves
$5.01\times$--$8.68\times$ speedups.
This behavior is consistent with the complexity discussion in
\Cref{tab:complexity-comparison}: ZKBoost pays heavily for depth-sensitive
route/path-consistency checks, whereas \textsc{Terrae} avoids proving explicit
root-to-leaf path objects and instead works with aggregated algebraic
constraints.  As a result, increasing depth has a much weaker effect on
\textsc{Terrae}, especially for inference.  The same effect appears in memory:
for inference, \textsc{Terrae} stays near $1.25$--$1.28$\,GB throughout the depth
sweep, while ZKBoost rises to $16.27$\,GB at $h=4$.

\paragraph{Scaling with the number of samples.}
When varying $M$, both systems again scale approximately linearly, reflecting
the fact that sample-dependent aggregation remains a dominant cost in both
designs.  Still, \textsc{Terrae} preserves a clear advantage throughout the
sweep: $4.46\times$--$5.09\times$ for training on completed baseline runs and
$5.17\times$--$6.05\times$ for inference.  At the largest tested point,
$M=4096$, ZKBoost times out during training, while \textsc{Terrae} completes in
$546.71$\,s; for inference, \textsc{Terrae} reduces proving time from
$876.30$\,s to $144.80$\,s.  This indicates that the gains of
\textsc{Terrae} are maintained even when the workload is dominated by
sample-level processing.  Memory usage also increases with $M$, but the gap
remains visible at the largest tested point: for $M=4096$, \textsc{Terrae} uses
$54.12$\,GB for training and $9.58$\,GB for inference, while ZKBoost times out
in training and uses $32.69$\,GB for inference.

Overall, \Cref{tab:three-dim-scaling} shows a consistent pattern: \textsc{Terrae}
tracks the natural linear growth in $T$ and $M$, while substantially reducing
the constant factors, and it is markedly more robust to increasing depth.  The
depth sweep is particularly important in practice, since deeper trees are
exactly the setting in which path-based certification becomes expensive.  For
both systems, proof size remains essentially constant across these experiments,
and verifier time stays small; the main practical differences are therefore
prover time and peak prover memory, both of which favor \textsc{Terrae}.

\subsection{Histogram Aggregation Microbenchmark}
\label{subsec:histogram-microbenchmark}

We isolate the histogram primitive from the full GBDT pipeline to measure the
benefit of replacing a generic RAM-style update proof with the specialized
aggregation argument of \textsc{Terrae}. The benchmark relation is the
following. Given
$\mathbf{ind}\in[m]^n$, $\mathbf{u}\in\mathbb{F}^n$, and
$\mathbf{v}\in\mathbb{F}^m$, the prover certifies that
\[
    \mathbf{v}[i]
    =
    \sum_{j:\mathbf{ind}[j]=i}\mathbf{u}[j],
    \quad i\in[m].
\]
This is the core operation used when sample-wise residual updates are converted
into per-node and per-bin histograms during training.

\paragraph{Baseline.}
The RAM baseline implements the same relation as a sequence of memory updates.
The memory array has $m$ addresses and is initialized to zero. For each
$j\in[n]$, the prover reads address $\mathbf{ind}[j]$, adds $\mathbf{u}[j]$,
writes the new value back, and finally proves that the resulting memory
equals $\mathbf{v}$.
Such baseline histogram proof is adopted in \cite{sparrow24,zkboost26}.

In contrast, \textsc{Terrae} proves the
same relation through the randomized aggregate identity in
\Cref{eq:histogram-aggregation-power-sum}, plus the power-table and lookup
checks needed to bind $\alpha^{\mathbf{ind}[j]}$.

\Cref{tab:histogram-n-scale} fixes $m=2^{15}$ and scales the number of updates
$n$. \Cref{tab:histogram-m-scale} fixes $n=2^{17}$ and scales the address
space $m$.

\input{tables/histogram_n_scale}

\input{tables/histogram_m_scale}

\paragraph{Scaling with the number of updates.}
When $m$ is fixed, the growth rates of both approaches approximate $2$
when $n$ becomes large and dominates the proof cost. In
\Cref{tab:histogram-n-scale}, increasing $n$ from $2^{15}$ to $2^{19}$
increases the prover time of our protocol from $8.91$\,s to $73.92$\,s, while
the RAM baseline grows from $56.27$\,s to $428.57$\,s. The specialized
histogram proof only adds one weighted contribution
$\mathbf{u}[j]\alpha^{\mathbf{ind}[j]}$ per update, plus a lookup into the
power table. The RAM baseline must additionally prove read-write consistency
for the evolving memory state. As a result, our protocol reduces prover time
by $6.31\times$ at $n=2^{15}$, $5.70\times$ at $n=2^{17}$, and
$5.80\times$ at $n=2^{19}$.
\textcolor{red}{The verifier time of both protocols is constant across this
sweep.  Our protocol also enjoys a smaller constant proof size: the specialized
proof is $2.84$\,KB, compared with about $8.0$\,KB for the RAM baseline.}

\paragraph{Scaling with the address space.}
Both protocols approach a $2\times$ growth rate
when $m=2^{17}$ and the address space becomes the dominant term. This is
the expected linear dependence on the number of histogram buckets. The RAM
baseline also grows with $m$, because the final memory image has more cells and
the consistency proof must bind a larger address universe. Its dominant cost,
however, is still the read-write trace over $n$ updates, so the slope in this
sweep is smaller than in the $n$-scale experiment until the memory image becomes
large. This leads to a widening prover-time gap: our protocol is
$5.76\times$ faster at $m=2^{14}$, $5.70\times$ faster at $m=2^{15}$,
$6.13\times$ faster at $m=2^{16}$, $10.95\times$ faster at $m=2^{17}$, and
$14.75\times$ faster at $m=2^{18}$.
\textcolor{red}{The verifier time remains constant for both protocols.  The
proof size also remains constant: $2.84$\,KB for \textsc{Terrae} and about
$8.0$\,KB for RAM, confirming that the histogram argument does not need to
expose a growing memory transcript.}

Overall, the histogram-proof benchmark supports the design choice used in \textsc{Terrae}.
This also supports and partly explains why the full-system gains in
\Cref{tab:three-dim-scaling} persist in the training pipeline, where histogram
construction is one of the dominant costs.
