\section{Open Science}\label{sec:open-science}
For reproducibility, we release an anonymous artifact containing the
prototype implementation, scripts for running the experiments, and the data
processing pipeline used in the evaluation. The artifact also includes
instructions for rebuilding the binaries and reproducing the main benchmark
tables reported in the paper. The anonymous repository is available at
\url{https://anonymous.4open.science/r/ZK-GBDT-F601}. We will
maintain the repository throughout the review period and, if the paper is
accepted, we plan to de-anonymize the repository and provide a permanent public
release.

\section{Ethical Considerations}
This work is a systems and cryptography study on zero-knowledge proofs for GBDT
training and inference. Its goal is to improve the verifiability and privacy of
machine-learning pipelines, especially in settings where a client wishes to
check that a model or prediction was produced correctly without revealing the
underlying data or model parameters. We do not introduce new datasets involving
human subjects, and our experiments use standard benchmark data and
machine-learning workloads. As with many privacy-enhancing and integrity
technologies, the techniques developed here could in principle be used in both
benign and harmful settings. In our view, the primary effect of this work is to
improve accountability in outsourced or untrusted ML services by enabling
auditable computation without exposing sensitive data. We are not aware of any
immediate ethical concerns beyond those already present in the underlying use of
ML models and cryptographic tools.
