\section{Soundness Analysis}\label{app:soundness}

In this appendix we analyse the soundness loss of the three
main proofs, i.e. proof of inference, proof of forest well-formedness, and proof of training. 

\paragraph{Primitive errors.}
We write $\epsilon_{\mathsf{KZG}}$ for the binding/opening soundness error of
the KZG backend, including the final batched openings.  We write
$\epsilon_{\mathsf{lookup}}(Q,T)$ for the soundness error of the imported lookup
argument on query length $Q$ and table length $T$.  Range checks are direct
lookups into a table of size $\mathcal{B}$, so they are included in the same
lookup term.  Maximum, argmax, fixed-point division, and quantization checks are
implemented by low-degree identities plus lookup checks, so their additional
loss is accounted for by the same terms.

Domain-lifting batching uses one random linear-combination challenge for each
batched family.  By \Cref{thm:domain-lifting-batching}, batching $m$ same-form
zero-check or sum-check claims contributes at most
\[
    \epsilon_{\mathsf{dl}}(m) \leq \frac{m-1}{|\mathbb{F}|}.
\]
Interleaving itself is deterministic and introduces no soundness loss.  When
interleaving is used to merge lookup families over distinct tables, the random
table shift in \Cref{app:interleaving-batching} contributes at most
\[
    \epsilon_{\mathsf{sep}}(s,L)
    \leq
    \frac{s^2L^2}{|\mathbb{F}|},
\]
where $s$ is the number of table families and $L$ upper bounds the table
lengths in the batch.  Finally, each histogram aggregation identity over an
address space of size $m$ contributes at most
\[
    \epsilon_{\mathsf{hist}}(m) \leq \frac{m-1}{|\mathbb{F}|},
\]
as shown in \Cref{subsec:overview-histogram}.

\paragraph{Inference.}
An accepting inference proof is sound unless one of the following events occurs.
First, a KZG commitment is opened inconsistently, contributing
$\epsilon_{\mathsf{KZG}}$.  Second, the leaf-fetch lookup accepts even though a
claimed row
\[
  \mathbf{1}\|\mathbf{ind\_tree}\|\mathbf{U}^{\leftarrow}\|
  \mathbf{U}^{\rightarrow}\|\mathbf{sc}
\]
does not appear in the committed forest table; this is bounded by the lookup
error plus the table-separation loss if several lookup families are batched.
Third, an interval containment check or final maximum-selection check accepts
incorrectly; these are bounded by the lookup errors for the corresponding
batches.  Fourth, a score aggregation identity fails but passes the batched
zero-check or sum-check; if $m_{\mathsf{inf},0}$ zero-check claims and
$m_{\mathsf{inf},\Sigma}$ sum-check claims are batched, this contributes
\[
    \epsilon_{\mathsf{dl}}(m_{\mathsf{inf},0})
    +
    \epsilon_{\mathsf{dl}}(m_{\mathsf{inf},\Sigma}).
\]
Thus the inference soundness error is bounded by
\[
\epsilon_{\mathsf{inf}}
\leq
\epsilon_{\mathsf{KZG}}
+\epsilon_{\mathsf{lookup}}^{\mathsf{inf}}
+\epsilon_{\mathsf{sep}}^{\mathsf{inf}}
+\epsilon_{\mathsf{dl}}(m_{\mathsf{inf},0})
+\epsilon_{\mathsf{dl}}(m_{\mathsf{inf},\Sigma}),
\]
which remains negligible for polynomially bounded batch and forest sizes and a sufficiently large field.

\paragraph{Forest well-formedness.}
The well-formedness proof consists of local algebraic constraints, lookup
checks for pointers and bounded fields, and one randomized root-uniqueness
argument.  The booleanity, no-ghost-node, parent-child, depth, hyper-box, and
feature-selection constraints are low-degree identities.  If
$m_{\mathsf{wf},0}$ such identities and $m_{\mathsf{wf},\Sigma}$ sum claims are
batched, their loss is
\[
    \epsilon_{\mathsf{dl}}(m_{\mathsf{wf},0})
    +
    \epsilon_{\mathsf{dl}}(m_{\mathsf{wf},\Sigma}).
\]
Pointer fetches, tree/depth fetches, and bounded-field checks contribute
$\epsilon_{\mathsf{lookup}}^{\mathsf{wf}}$,
and, when applicable, $\epsilon_{\mathsf{sep}}^{\mathsf{wf}}$.  The
root-uniqueness check is a histogram-style power-sum identity over the $TK+1$
tree identifiers, where identifier $0$ is the dummy value.  If the committed
root-count vector is incorrect, it passes with probability at most
$TK/|\mathbb{F}|$.  Therefore
\[
\epsilon_{\mathsf{wf}}
\leq
\epsilon_{\mathsf{KZG}}
+\epsilon_{\mathsf{lookup}}^{\mathsf{wf}}
+\epsilon_{\mathsf{sep}}^{\mathsf{wf}}
+\epsilon_{\mathsf{dl}}(m_{\mathsf{wf},0})
+\epsilon_{\mathsf{dl}}(m_{\mathsf{wf},\Sigma})
+\frac{TK}{|\mathbb{F}|}.
\]

\paragraph{Training.}
The training proof includes the well-formedness proof for the committed forest
and then checks that the auxiliary training trace realizes the public
quantized trainer.  Its additional soundness loss comes from five sources.
First, routing the training samples to leaves uses the same lookup and
low-degree checks as batched inference, now over $TKM$ routed rows.  Second,
the boosting-state updates, score aggregation, prefix sums, leaf-value
equations, and split-score equations are low-degree or sum-check relations and
are covered by domain-lifting batching.  Third, fixed-point division,
quantization, exponential-table, bounded-field, and maximum/argmax checks are
accounted for by lookup errors.  Fourth, histogram construction uses randomized
power-sum identities.  If the proof contains histogram checks over address
spaces $m_1,\ldots,m_H$, their total loss is
\[
    \sum_{\ell=1}^{H}\epsilon_{\mathsf{hist}}(m_\ell)
    \leq
    \sum_{\ell=1}^{H}\frac{m_\ell-1}{|\mathbb{F}|}.
\]
In the main GBDT histogram, the natural address space is
$N_{\mathrm{tot}}dB$, corresponding to node-feature-bin buckets.  Fifth, all
batched openings and commitments contribute the backend term
$\epsilon_{\mathsf{KZG}}$.

Let $m_{\mathsf{tr},0}$ and $m_{\mathsf{tr},\Sigma}$ be the numbers of
training zero-check and sum-check claims after grouping by same form, and let
the superscripted $\epsilon$ terms denote the lookup and table-separation losses
of the training batches.  A union bound gives
\[
\begin{aligned}
\epsilon_{\mathsf{tr}}
\leq{}&
\epsilon_{\mathsf{wf}}
+\epsilon_{\mathsf{KZG}}
+\epsilon_{\mathsf{lookup}}^{\mathsf{tr}}
+\epsilon_{\mathsf{sep}}^{\mathsf{tr}} \\
&+\epsilon_{\mathsf{dl}}(m_{\mathsf{tr},0})
+\epsilon_{\mathsf{dl}}(m_{\mathsf{tr},\Sigma})
+\sum_{\ell=1}^{H}\frac{m_\ell-1}{|\mathbb{F}|}.
\end{aligned}
\]
Consequently, except with this probability, an accepting training proof binds
the committed dataset, forest, and training trace to a valid execution of the
public quantized trainer.
Since all sizes are polynomially bounded,
the training soundness error is negligible for a sufficiently large field.
